---
title: "Catch only projection for petrale sole"
author: "Ian G. Taylor" 
affiliation: 'NOAA Fisheries Northwest Fisheries Science Center, 2725 Montlake Blvd E, Seattle, WA 98112'
date: today
format: 
  pdf:
    template: catch_only_projection_template.tex
params:
  species: "petrale sole"
  assess_year: 2023
  current_year: 2025
  final_proj_year: 2036
  category: 1
  sigma: 0.50
  p_star: 0.45
  spawn_output_decimals: 2
  dir_original: "../models/2023.a034.920_base_Pstar45"
  dir_new: "2025_standard_3.30.24_july15"
  dir_new2: "2025_alternative1_3.30.24_july15"
  dir_new3: "2025_Nov20"
  spawn_output_units: "(trillions of eggs)"
---


```{r}
#| label: parameters
#| echo: false
#| warning: false
#| message: false
species <- params$species
Species <- stringr::str_to_sentence(params$species)
assess_year <- params$assess_year
current_year <- params$current_year
final_proj_year <- params$final_proj_year
category <- params$category
sigma <- params$sigma
p_star <- params$p_star
spawn_output_decimals <- params$spawn_output_decimals

# original model
model_original <- r4ss::SS_output(
    dir = params$dir_original,
    verbose = FALSE,
    printstats = FALSE,
    hidewarn = TRUE
)

# catch-only projection
model <- r4ss::SS_output(
    dir = params$dir_new, # this is the directory where I ran the standard Catch-only projection model
    verbose = FALSE,
    printstats = FALSE,
    hidewarn = TRUE
)

# catch-only projection Alternative 1
model2 <- r4ss::SS_output(
    dir = params$dir_new2, # this is the directory where I ran the standard Catch-only projection model
    verbose = FALSE,
    printstats = FALSE,
    hidewarn = TRUE
)

# catch-only projection Alternative 2 (Nov 2025)
model3 <- r4ss::SS_output(
    dir = params$dir_new3,
    verbose = FALSE,
    printstats = FALSE,
    hidewarn = TRUE
)

table_caption <- glue::glue("Original {assess_year} projection and new catch-only projection for OFLs (mt), buffer, ABCs (mt), actual & assumed catch (mt), ACLs (mt), spawning output {params$spawn_output_units}, and stock status (fraction of unfished spawning output) given the removals. The gray shading indicates values associated with the original {assess_year} assessment projection. In both the original model and the new projection, the removals for years beyond the values in the 'Actual & Assumed Removals' columns are the ACL values estimated from that model.")
```


This document details a catch-only projection for `r species` (Table 1 below). The most recent assessment for `r species` was conducted in `r assess_year`. The projection presented here is based on an previous catch-only projection described in Agenda Item G.8 Supplemental Attachment 3 September 2025 https://www.pcouncil.org/documents/2025/09/g-8-supplemental-attachment-3-catch-only-projection-for-petrale-sole.pdf/. The catches for 2023--2026 match Table 2 in that previous catch-only projection up through 2026 but differ for 2027 and beyond. The catches in each year in this new projection are based on the following:

  - 2023--2024: observed removals as provided by the Groundfish Management Team (GMT)
  - 2025: removals set equal to mortality projections provided by the GMT for each fleet in the model which includes an assumed attainment of 92% estimated by the GMT.
  - 2026: removals set at 2290 mt based on the ACL from a catch-only projection through 2025 with an assumption of 92% attainment (2489 * 0.92 = 2290)
  - 2027--2028: removals set equal to the 2026 ACL with an assumption of 100% attainment (2489 mt)
  - 2029--2036: removals were set equal to the projected Annual Catch Limit (ACL) based on a category `r category` time-varying $\sigma$ of `r sigma` and a P* value of `r p_star`.
 
```{r}
#| label: calc-table
#| echo: false
#| warning: false
#| message: false

# define ranges of years to get values
all_years <- assess_year:final_proj_year
proj_years <- (current_year + 2):final_proj_year
proj_years2 <- (current_year + 1):final_proj_year
catch_years <- assess_year:(current_year + 1)
catch_years <- 2025:2026
catch_years2 <- assess_year:(current_year)
diff_years <- current_year - assess_year

# function to get model3 outputs
catch_only_projection_table <- function(model3, all_years, proj_years, catch_years) {
    # get realized catch from model3 output
    catch <- model3$derived_quants |>
        dplyr::filter(Label %in% paste0("ForeCatch_", catch_years)) |>
        dplyr::reframe(
            Year = stringr::str_extract(Label, "\\d+") |> as.numeric(),
            Catch = Value
        )
    ofl <- model3$derived_quants |>
        dplyr::filter(Label %in% paste0("OFLCatch_", proj_years)) |>
        dplyr::reframe(
            Year = stringr::str_extract(Label, "\\d+") |> as.numeric(),
            OFL = Value
        )
    acl <- model3$derived_quants |>
        dplyr::filter(Label %in% paste0("ForeCatch_", proj_years)) |>
        dplyr::reframe(
            Year = stringr::str_extract(Label, "\\d+") |> as.numeric(),
            ACL = Value
        )
    sb <- model3$derived_quants |>
        dplyr::filter(Label %in% paste0("SSB_", all_years)) |>
        dplyr::reframe(
            Year = stringr::str_extract(Label, "\\d+") |> as.numeric(),
            SB = Value
        )
    status <- model3$derived_quants |>
        dplyr::filter(Label %in% paste0("Bratio_", all_years)) |>
        dplyr::reframe(
            Year = stringr::str_extract(Label, "\\d+") |> as.numeric(),
            status = round(Value, 3)
        )

    # get buffer from PEPtools function
    Buffer <- PEPtools::get_buffer(
        years = all_years, sigma = sigma, pstar = p_star,
        verbose = FALSE
    ) |>
        dplyr::rename(Year = year, Buffer = buffer) |>
        dplyr::filter(Year %in% proj_years)

    # combine all the tables above (joined by year)
    table <- purrr::reduce(
        list(ofl, Buffer, acl, sb, status, catch),
        ~ dplyr::full_join(.x, .y, by = "Year")
    ) |>
        dplyr::arrange(Year) |>
        dplyr::relocate(Catch, .after = Year) |>
        dplyr::mutate(
            Buffer_from_ratio = dplyr::case_when(
                status >= model3$btarg ~ round(ACL / OFL, 3),
                .default = NA # ratio doesn't work for years with 40-10 adjustment
            ),
            .after = Buffer
        ) |>
        dplyr::mutate(
            ABC = Buffer * OFL,
            .after = Buffer_from_ratio
        ) |>
        dplyr::rename(
            `Actual & Assumed Removals` = Catch,
            `Stock Status` = status,
            `Spawning Output` = SB,
        )
}

# function to apply gt formatting to the table
format_table <- function(table) {
    table |>
        gt::gt() |>
        gt::fmt_number(
            columns = tidyselect::starts_with("Actual") | tidyselect::starts_with("OFL") | tidyselect::starts_with("ABC") | tidyselect::starts_with("ACL"),
            decimals = 0
        ) |>
        gt::fmt_number(
            columns = tidyselect::contains("Spawning"),
            decimals = spawn_output_decimals
        ) |>
        gt::tab_options(
            table.font.size = 11, # reduced font size
            latex.use_longtable = TRUE
        ) |>
        gt::sub_missing(
            columns = tidyselect::everything(),
            missing_text = "---"
        ) |>
        gt::cols_align(
            align = "center"
        ) |>
        gt::cols_width(
            everything() ~ px(65)
        ) |>
        gt::data_color(
            columns = tidyselect::contains("HCR"),
            palette = "gray90",
            na_color = "gray90"
        ) |>
        gt::as_latex()
}

table <- catch_only_projection_table(model3,
    all_years = all_years, proj_years = proj_years, catch_years = catch_years
) |>
    dplyr::select(-`Spawning Output`, -`Actual & Assumed Removals`, -Buffer) |>
    dplyr::mutate( # Buffer = Buffer_from_ratio, # overwrite buffer from PEPtools::get_buffer
        ABC = ACL
    ) |> # overwrite externally calculated ABC
    dplyr::rename_with(~ paste(.x, "(yyyy)"), .cols = -Year)

table_new2 <- catch_only_projection_table(model2,
    all_years = head(all_years, 12), proj_years = proj_years - diff_years, catch_years = catch_years - diff_years
) |>
    dplyr::select(-`Spawning Output`)


```

```{r}
#| label: warning-about-buffer
#| echo: false
#| warning: true
#| message: true

# combine the two tables

wide_table3 <- dplyr::full_join(
    table,
    table_new2,
    by = "Year"
) |>
    dplyr::select(
        Year,
        `OFL (yyyy)`, OFL,
        `Buffer`,
        `ABC (yyyy)`, ABC,
        `Actual & Assumed Removals`,
        `ACL (yyyy)`, ACL,
        `Stock Status (yyyy)`, `Stock Status`
    ) |>
    dplyr::rename_with(.fn = function(x) {
        gsub("yyyy", "alt HCR", x)
    })

```

::: {.landscape}
```{r}
#| label: tbl-proj3
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| tbl-cap: !expr table_caption
format_table(wide_table3)
```

:::