```{r executive, echo = FALSE}
executive <- list()
executive[["stock"]] <- paste0("This assessment reports the status of ",
  spp, " (_", spp.sci, "_) off the ", coast, 
  " coast using data through 2022.")
```

# One-page summary{-}

- test
- test
- test

\clearpage

# Executive summary{-}

## Stock{-}

`r executive[["stock"]]` While `r spp` are modeled as a single stock, the spatial aspects of the coast-wide population are addressed through geographic separation of data sources/fleets where possible. There is currently no genetic evidence suggesting distinct biological stocks of `r spp` off the U.S. coast. The limited tagging data available to describe adult movement suggests that `r spp` may have some homing ability for deep water spawning sites but also have the ability to move long distances between spawning sites.


## Catches{-}

`r Spp` has  consistently been the most commercially valuable flatfish targeted in the California Current Ecosystem. It has been caught in the flatfish fishery off the U.S. Pacific coast since the late 19th century, nearly exclusively by trawl fleets; non-trawl gears contribute only a small fraction of the catches across all years. 

The earliest catches of `r spp` are reported in 1876 in California and 1884 in Oregon.  `r Spp` were lightly exploited during the early 1900s, but new gear technology in the 1930s allowed trawling on new grounds and the fishery expanded to greater depths and to Oregon and Washington waters, resulting in larger landings. The `r spp` catches further increased during World War II in response to increased demands. Also, during the "vitamin A rush" in the late 1930s and 1940s it was found that `r spp` has high levels, which contributed to increased catches of this species as well. By the 1950s, the fishery was well developed with the stock showing declines in biomass and catches (Figures \ref{fig:es-catch} and \ref{fig:es-ssb}). Also in the 1950s, winter spawning grounds at deeper depths with dense concentrations of `r spp` were discovered, and catches increased accordingly. The rate of decline in spawning biomass accelerated through the 1970s reaching minimums generally around or below 10% of the unexploited levels during the 1980s through the early 2000s (Figure \ref{fig:es-depl}). Recent annual catches between 1981-2022 range between 803 and 3060 mt per year and the most recent year landings are shown in Table \ref{tab:removalsES}. 

`r Spp` are a desirable market species and discarding has historically been low. 

In this assessment, fishery removals have been divided between two fleets: 1) North and 2) South. Landings for the North fleet are defined as fish landed in Washington and Oregon ports. Landings for the South fleet are defined as fish landed in California ports.

\clearpage

\input{tex_tables/a_Catches_ES.tex}

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc_relative, "plots", "catch2 landings stacked.png"), 
caption = "Landings by fleet used in the base model where catches in metric tons by fleet are stacked",
label = 'es-catch')
```

\clearpage

## Data and assessment{-}

This assessment uses the stock assessment framework Stock Synthesis 
```{r es-modelversion, echo=FALSE} 
if (exists("model")) {
  strsplit(model$SS_version,";")[[1]][1]
}
```
(SS3).

The modeling period begins in 1876, assuming an unfished equilibrium state of the stock in 1875. The assessment treats females and males separately due to differences in biology and life history parameters between sexes, and the coastwide population is modeled allowing separate growth and mortality parameters for each sex (a two-sex model). The model also allows for differences in selectivity between sexes. 

Types of data that inform the model include catch, length and age frequency data from two commercial fleets, that include 1) North and 2) South. Biological data are derived from both port and on-board observer sampling programs. The National Marine Fisheries Service (NMFS) \gls{s-tri}  (1980, 1983, 1986, 1989, 1992, 1995, 1998, 2001, and 2004) and the  \Gls{s-wcgbt} (2003-2022) relative biomass indices and biological sampling provide fishery independent information on relative trend and demographics of the `r spp` stock.

The definition of fishing fleets changed in this assessment relative to the 2019 assessment. Previously, North and South fleets were additionally divided into seasonal fisheries (Winter and Summer), with fishing year starting in November of previous year. In this assessment, Winter and Summer fisheries are combined within each fleet into annual fisheries with removals corresponding to calendar year.

Growth is assumed to follow the von Bertalanffy growth model, and the assessment explicitly estimates all parameters describing somatic growth. Natural mortality is estimated for both sexes using meta-analytical prior. Externally estimated life history parameters, including those defining the length-weight relationship, female fecundity and maturity schedule were revised for this assessment to incorporate new information. Recruitment dynamics are assumed to follow the Beverton-Holt stock-recruit function, and recruitment deviations are estimated. Stock-recruitment steepness is fixed at the value generated from meta-analytical study.

## Stock biomass and dynamics{-}

Replace text with
trends and current levels relative to virgin or historic levels and
description of uncertainty.
Include Table for last 10 years.
Include Figure with long-term estimates.

\input{tex_tables/b_SSB_ES.tex}

```{r, results = 'asis'}
if (!is.na(mod_base$log_det_hessian)) {
  # with Hessian
  add_figure(
  filein = file.path(mod_loc_relative, "plots",
  "ts7_Spawning_output_with_95_asymptotic_intervals_intervals.png"), 
  caption = "Estimated time series of spawning output (circles and line: median; light broken lines: 95 percent intervals) for the base model",
  label = 'es-ssb')
} else {
  # no uncertainty (model not converged or run with -nohess)
  add_figure(
  filein = file.path(mod_loc_relative, "plots",
  "ts7_Spawning_output.png"), 
  caption = "Estimated time series of spawning output for the base model",
  label = 'es-ssb')
}
```

```{r, results = 'asis'}
if (!is.na(mod_base$log_det_hessian)) {
  # with Hessian
  add_figure(
  filein = file.path(mod_loc_relative, "plots", "ts9_Relative_spawning_output_intervals.png"), 
  caption = "Estimated time series of fraction of unfished spawning output (circles and line: median; light broken lines: 95 percent intervals) for the base model",
  label = 'es-depl')
} else {
  # no uncertainty (model not converged or run with -nohess)
  add_figure(
  filein = file.path(mod_loc_relative, "plots", "ts9_Relative_spawning_output.png"), 
  caption = "Estimated time series of fraction of unfished spawning output for the base model",
  label = 'es-depl')
}
```

\clearpage

## Recruitment{-}

Replace text with
trends and current levels relative to virgin or historic levels and
description of uncertainty.
Include Table for last 10 years.
Include Figure with long-term estimates.


\input{tex_tables/c_Recr_ES.tex}

```{r, results = 'asis'}
if (!is.na(mod_base$log_det_hessian)) {
  # with Hessian
  add_figure(
  filein = file.path(mod_loc_relative, "plots", "ts11_Age-0_recruits_(1000s)_with_95_asymptotic_intervals.png"), 
  caption = "Estimated time series of age-0 recruits (1000s) for the base model with 95 percent intervals",
  label = 'es-recruits')
} else {
  # no uncertainty (model not converged or run with -nohess)
  add_figure(
  filein = file.path(mod_loc_relative, "plots", "ts11_Age-0_recruits_(1000s).png"), 
  caption = "Estimated time series of age-0 recruits (1000s) for the base model",
  label = 'es-recruits')
}
```

```{r, results = 'asis'}
if (!is.na(mod_base$log_det_hessian)) {
  # with Hessian
  add_figure(
  filein = file.path(mod_loc_relative, "plots", "recdevs2_withbars.png"), 
  caption = "Estimated time series of recruitment deviations",
  label = 'es-rec-devs')
} else {
  # no uncertainty (model not converged or run with -nohess)
  add_figure(
  filein = file.path(mod_loc_relative, "plots", "recdevs1_points.png"), 
  caption = "Estimated time series of recruitment deviations",
  label = 'es-rec-devs')
}
```

\clearpage
## Exploitation status{-}

Replace text with
total catch divided by exploitable biomass or SPR harvest rate.
Include Table for last 10 years.
Include Figure with trend in f relative to target vs. trend in biomass relative to the target.

\input{tex_tables/d_SPR_ES.tex}

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc_relative, "plots", "SPR2_minusSPRseries.png"), 
caption = "Estimated 1 - relative spawning ratio (SPR) by year for the base model. The management target is plotted as a red horizontal line and values above this reflect harvest in excess of the proxy harvest rate",
label = 'es-1-spr')
```

\clearpage

## Ecosystem considerations{-}

Over the past several years, progress has been done in understanding how large-scale climate forcing drives regional changes and impacts density-independent survival during `r spp` early  life  stages. Haltuch et al. (2019) examined the relationship between `r spp` recruitment and oceanographic drivers based on model output from a Regional Ocean Modeling System (ROMS) model for the California Current Ecosystem (Neveu et al. 2016). The results suggested that ROMS output might be useful as the basis for an environmental index of recruitment for `r spp`, to allow for better model precision and near-term forecasting. However, the ROMS model used by Haltuch et al. (2019) was consistent in structure and inputs for 1980-2010. From 2011 forward, ROMS outputs exhibit distinct discontinuities with 1980-2010 period, showing changes in scale and trend across the 2010/2011 boundary. 

Due to discontinuity in ROMS models before and after 2011, index developed by Haltuch et al. (2019) could not be used for this stock assessment. However, an alternative oceanographic model products by Copernicus Marine Environment Monitoring Service (CMEMS) (https://marine.copernicus.eu/) and Mercator Ocean International (MOI) (https://www.mercator-ocean.eu/) were investigated to test if this modelling framework could be used to produce an environmental index of `r spp` recruitment. 

Appendix 1 of this report describes most recent efforts in developing environmental index of `r spp` recruitment based on CMEMS products. Appendix 1 also discusses potential reasons for why ROMS time series changed over the 2010/2011 boundary and how those changes might impact the ROMS-recruitment model from Haltuch et al. (2019).

In this assessment, we explore the impact of current CMEMS based environmental index of `r spp` recruitment via sensitivity analysis. 

## Reference points{-}

Replace text with
management targets and definition of overfishing, including the harvest rate that brings the stock to equilibrium at $B_{40\%}$, i.e., the $B_{MSY}$ proxy and the equilibrium stock size that results from fishing at the default harvest rate, i.e., the $F_{MSY}$ proxy.
Include Table of estimated reference points for ssb, SPR, exploitation rate, and yield based on SSB proxy for MSY, SPR proxy for MSY, and estimated MSY values.

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc_relative, "plots", "SPR4_phase.png"), 
caption = "Phase plot of estimated 1-SPR versus fraction unfished for the base model",
label = 'es-phase')
```

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc_relative, "plots", "yield2_yield_curve_with_refpoints.png"), 
caption = "Equilibrium yield curve for the base case model. Values are based on the 2020
fishery selectivities and with steepness fixed at 0.80",
label = 'es-yield')
```

\input{tex_tables/e_ReferencePoints_ES.tex}

\clearpage

## Management performance{-}

In the last ten years total dead catches (including estimated dead discards) 
have been below the annual catch limit, although the attainment has greater than
90\% in many of those years.

```{r, tables-vi, results = 'asis'}
table_vi <- read.csv("../tables/table_vi.csv")
sa4ss::table_format(x = table_vi,
                 caption = paste("Recent trend in the overfishing limits (OFLs),",
                 "the acceptable biological catches (ABCs),",
                 "the annual catch limits (ACLs), the total landings, ",
                 "and total mortality (mt).",
                 "The 'Total dead' values may differ from the GEMM estimates because",
                 "the discard mortality is computed within the model based on estimated retention curves."),
                 label = 'table-vi'
                 )
```

## Unresolved problems and major uncertainties{-}

Replace text with
any special issues that complicate scientific assessment, questions about the best model scenario, etc.

## Decision table and projections{-}

Replace text with
projected yields (OFL, ABC, and ACL), spawning biomass, and stock depletion levels for each year.
OFL calculations should be based on the assumption that future catches equal ABCs and not OFLs.

## Scientific uncertainty{-}

Replace text with
the sigma value and the basis for its calculation.

## Research and data needs{-}

Replace text with
information gaps that seriously impede the stock assessment.
