```{r executive, echo = FALSE}
executive <- list()
executive[["stock"]] <- paste0("This assessment reports the status of ",
  spp, " (_", spp.sci, "_) off the ", coast, 
  " coast using data through 2022.")
```

# Executive summary{-}

## Stock{-}

`r executive[["stock"]]`

## Catches{-}

Adding text here for example.
trends and current levels.
Include Table for last 10 years.
Include Figure with long-term estimates.

\clearpage

\input{tex_tables/a_Catches_ES.tex}

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc_relative, "plots", "catch2 landings stacked.png"), 
caption = "Landings by fleet used in the base model where catches in metric tons by fleet are stacked",
label = 'es-catch')
```

\clearpage

## Data and assessment{-}

This assessment uses the stock assessment framework
Stock Synthesis 
```{r es-modelversion, echo=FALSE} 
if (exists("model")) {
  strsplit(model$SS_version,";")[[1]][1]
}
```
(SS3).

Replace text with
date of last assessment,
type of assessment model,
data available,
new information, and
information lacking.

## Stock biomass and dynamics{-}

Replace text with
trends and current levels relative to virgin or historic levels and
description of uncertainty.
Include Table for last 10 years.
Include Figure with long-term estimates.

\input{tex_tables/b_SSB_ES.tex}

```{r, results = 'asis'}
if (!is.na(mod_base$log_det_hessian)) {
  # with Hessian
  add_figure(
  filein = file.path(mod_loc_relative, "plots",
  "ts7_Spawning_output_with_95_asymptotic_intervals_intervals.png"), 
  caption = "Estimated time series of spawning output (circles and line: median; light broken lines: 95 percent intervals) for the base model",
  label = 'es-ssb')
} else {
  # no uncertainty (model not converged or run with -nohess)
  add_figure(
  filein = file.path(mod_loc_relative, "plots",
  "ts7_Spawning_output.png"), 
  caption = "Estimated time series of spawning output for the base model",
  label = 'es-ssb')
}
```

```{r, results = 'asis'}
if (!is.na(mod_base$log_det_hessian)) {
  # with Hessian
  add_figure(
  filein = file.path(mod_loc_relative, "plots", "ts9_Relative_spawning_output_intervals.png"), 
  caption = "Estimated time series of fraction of unfished spawning output (circles and line: median; light broken lines: 95 percent intervals) for the base model",
  label = 'es-depl')
} else {
  # no uncertainty (model not converged or run with -nohess)
  add_figure(
  filein = file.path(mod_loc_relative, "plots", "ts9_Relative_spawning_output.png"), 
  caption = "Estimated time series of fraction of unfished spawning output for the base model",
  label = 'es-depl')
}
```

\clearpage

## Recruitment{-}

Replace text with
trends and current levels relative to virgin or historic levels and
description of uncertainty.
Include Table for last 10 years.
Include Figure with long-term estimates.


\input{tex_tables/c_Recr_ES.tex}

```{r, results = 'asis'}
if (!is.na(mod_base$log_det_hessian)) {
  # with Hessian
  add_figure(
  filein = file.path(mod_loc_relative, "plots", "ts11_Age-0_recruits_(1000s)_with_95_asymptotic_intervals.png"), 
  caption = "Estimated time series of age-0 recruits (1000s) for the base model with 95 percent intervals",
  label = 'es-recruits')
} else {
  # no uncertainty (model not converged or run with -nohess)
  add_figure(
  filein = file.path(mod_loc_relative, "plots", "ts11_Age-0_recruits_(1000s).png"), 
  caption = "Estimated time series of age-0 recruits (1000s) for the base model",
  label = 'es-recruits')
}
```

```{r, results = 'asis'}
if (!is.na(mod_base$log_det_hessian)) {
  # with Hessian
  add_figure(
  filein = file.path(mod_loc_relative, "plots", "recdevs2_withbars.png"), 
  caption = "Estimated time series of recruitment deviations",
  label = 'es-rec-devs')
} else {
  # no uncertainty (model not converged or run with -nohess)
  add_figure(
  filein = file.path(mod_loc_relative, "plots", "recdevs1_points.png"), 
  caption = "Estimated time series of recruitment deviations",
  label = 'es-rec-devs')
}
```

\clearpage
## Exploitation status{-}

Replace text with
total catch divided by exploitable biomass or SPR harvest rate.
Include Table for last 10 years.
Include Figure with trend in f relative to target vs. trend in biomass relative to the target.

\input{tex_tables/d_SPR_ES.tex}

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc_relative, "plots", "SPR2_minusSPRseries.png"), 
caption = "Estimated 1 - relative spawning ratio (SPR) by year for the base model. The management target is plotted as a red horizontal line and values above this reflect harvest in excess of the proxy harvest rate",
label = 'es-1-spr')
```

\clearpage

## Ecosystem considerations{-}

Replace text with
a summary of reviewed environmental and ecosystem factors that appear to be correlated with stock dynamics.
These may include variability in they physical environment, habitat, competitors, prey, or predators that directly or indirectly affects the stock's status, vital rates (growth, survival, productivity/recruitment) or range and distribution.
Note which, if any, ecosystem factors are used in the assessment and how (e.g., as background information, in data preparations, as data inputs, in decisions about model structure).

## Reference points{-}

Replace text with
management targets and definition of overfishing, including the harvest rate that brings the stock to equilibrium at $B_{40\%}$, i.e., the $B_{MSY}$ proxy and the equilibrium stock size that results from fishing at the default harvest rate, i.e., the $F_{MSY}$ proxy.
Include Table of estimated reference points for ssb, SPR, exploitation rate, and yield based on SSB proxy for MSY, SPR proxy for MSY, and estimated MSY values.

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc_relative, "plots", "SPR4_phase.png"), 
caption = "Phase plot of estimated 1-SPR versus fraction unfished for the base model",
label = 'es-phase')
```

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc_relative, "plots", "yield2_yield_curve_with_refpoints.png"), 
caption = "Equilibrium yield curve for the base case model. Values are based on the 2020
fishery selectivities and with steepness fixed at 0.80",
label = 'es-yield')
```

\input{tex_tables/e_ReferencePoints_ES.tex}

\clearpage

## Management performance{-}

In the last ten years total dead catches (including estimated dead discards) 
have been below the annual catch limit, although the attainment has greater than
90\% in many of those years.

```{r, tables-vi, results = 'asis'}
table_vi <- read.csv("../tables/table_vi.csv")
sa4ss::table_format(x = table_vi,
                 caption = paste("Recent trend in the overfishing limits (OFLs),",
                 "the acceptable biological catches (ABCs),",
                 "the annual catch limits (ACLs), the total landings, ",
                 "and total mortality (mt).",
                 "The 'Total dead' values may differ from the GEMM estimates because",
                 "the discard mortality is computed within the model based on estimated retention curves."),
                 label = 'table-vi'
                 )
```

## Unresolved problems and major uncertainties{-}

Replace text with
any special issues that complicate scientific assessment, questions about the best model scenario, etc.

## Decision table and projections{-}

Replace text with
projected yields (OFL, ABC, and ACL), spawning biomass, and stock depletion levels for each year.
OFL calculations should be based on the assumption that future catches equal ABCs and not OFLs.

## Scientific uncertainty{-}

Replace text with
the sigma value and the basis for its calculation.

## Research and data needs{-}

Replace text with
information gaps that seriously impede the stock assessment.
