```{r executive, echo = FALSE}
executive <- list()
executive[["stock"]] <- paste0("This assessment reports the status of ",
  spp, " (_", spp.sci, "_) off the U.S. West",
  " Coast using data through 2022.")
```

# One-page summary{-}

- executive[["stock"]]
- `r Spp` is a commercially value flatfish primarily caught using bottom trawl gear.
- test

\clearpage

# Executive summary{-}

## Stock{-}

`r executive[["stock"]]` While `r spp` are modeled as a single stock, the spatial aspects of the coastwide population are addressed through geographic separation of data sources/fleets where possible. There is currently no genetic evidence suggesting distinct biological stocks of `r spp` off the U.S. West Coast. The limited tagging data available to describe adult movement suggests that `r spp` may have some homing ability for deep water spawning sites but also have the ability to move long distances between spawning sites.


## Catches{-}

`r Spp` has  consistently been the most commercially valuable flatfish targeted in the California Current Ecosystem. It has been caught in the flatfish fishery off the U.S. West Coast since the late 19th century, nearly exclusively by trawl fleets; non-trawl gears contribute only a small fraction of the catches across all years. 

The earliest catches of `r spp` are reported in 1876 in California and 1884 in Oregon.  `r Spp` were lightly exploited during the early 1900s, but new gear technology in the 1930s allowed trawling on new grounds and the fishery expanded to greater depths and to Oregon and Washington waters, resulting in larger landings. The `r spp` catches further increased during World War II in response to increased demands. Also, during the "vitamin A rush" in the late 1930s and 1940s it was found that `r spp` has high levels, which contributed to increased catches of this species as well. By the 1950s, the fishery was well developed with the stock showing declines in biomass and catches (Figures \ref{fig:es-catch} and \ref{fig:es-ssb}). Also in the 1950s, winter spawning grounds at deeper depths with dense concentrations of `r spp` were discovered, and catches increased accordingly. The rate of decline in spawning biomass accelerated through the 1970s reaching minimums estimated to be generally around or below 10% of the unexploited levels during the 1980s through the early 2000s (Figure \ref{fig:es-depl}). Recent annual catches between 1981-2022 range between 803 and 3060 mt per year and the most recent landings are shown in Table \ref{tab:removalsES}. 

`r Spp` are a desirable market species and discarding has historically been low (less than 5%), with most of the discarding due to small sizes. 

In this assessment, fishery removals have been divided between two fleets: 1) North and 2) South. Landings for the North fleet are defined as fish landed in Washington and Oregon ports. Landings for the South fleet are defined as fish landed in California ports.

\clearpage

```{r, results = 'asis'}
tab <- read.csv(file.path(mod_loc_relative, "tables/b_SSB_ES.csv"))
sa4ss::table_format(tab,
    caption = "Estimated ageing error vectors used in the assessment model for true ages 0-30 read by the Cooperative Ageing Project (CAP) lab. The ages associated with 'CAP Surface Pre-1990' were excluded from the base model. Note that the population age bins extend to age 40, and the largest observed age is 32.",
    label = 'ageing-error-cap'
  )
```

\input{tex_tables/a_Catches_ES.tex}

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc_relative, "plots", "catch2 landings stacked.png"), 
caption = "Landings by fleet used in the base model where catches in metric tons by fleet are stacked",
label = 'es-catch')
```

\clearpage

## Data and assessment{-}

This assessment uses the stock assessment framework Stock Synthesis (SS3; (Methot and Wetzel 2013)) version 3.30.21.00 released 10 February, 2023.
The last full assessment of `r spp` was conducted in 2013 and most recent update assessment in 2019. 

The modeling period begins in 1876, assuming an unfished equilibrium state of the stock in 1875. The assessment treats females and males separately due to differences in biology and life history parameters between sexes, and the coastwide population is modeled allowing separate growth and mortality parameters for each sex (a two-sex model). The model also allows for differences in selectivity between sexes. 

Types of data that inform the model include catch, length and age frequency data from two commercial fleets, that include 1) North and 2) South. Biological data are derived from both port and on-board observer sampling programs. The \gls{s-tri}  (1980, 1983, 1986, 1989, 1992, 1995, 1998, 2001, and 2004) and the \gls{s-wcgbt} (2003-2019, 2021-2022) relative biomass indices and biological sampling provide fishery independent information on relative trend and demographics of the `r spp` stock.

The definition of fishing fleets changed in this assessment relative to the 2019 assessment. Previously, North and South fleets were additionally divided into seasonal fisheries (Winter and Summer), with fishing year starting in November of previous year. In this assessment, Winter and Summer fisheries are combined within each fleet into annual fisheries with removals corresponding to calendar year.

Growth is assumed to follow the von Bertalanffy growth model, and the assessment explicitly estimates all parameters describing somatic growth. Natural mortality is estimated for both sexes using a meta-analytical prior. Externally estimated life history parameters, including those defining the length-weight relationship, female fecundity and maturity schedule were revised for this assessment to incorporate new information. Recruitment dynamics are assumed to follow the Beverton-Holt stock-recruit function, and recruitment deviations are estimated. Stock-recruitment steepness is fixed at the value generated from meta-analytical study.

## Stock biomass and dynamics{-}

```{r, results = "asis", eval = TRUE, echo = FALSE}
   # calculate some values from the timeseries
   ts <- mod_base$timeseries
   smry_equ <- ts$Bio_smry[ts$Era == "VIRG"]
   smry_2023 <- ts$Bio_smry[ts$Yr == 2023]
   smry_min <- min(ts$Bio_smry)
   smry_minyr <- ts$Yr[ts$Bio_smry %in% smry_min]
   smry_max_recent <- max(ts$Bio_smry[ts$Yr %in% 2010:2020])
   smry_maxyr_recent <- ts$Yr[ts$Bio_smry %in% smry_max_recent]
   
   # Bratio values
   sprseries <- mod_base$sprseries
   Bratio_2023 <- sprseries$Deplete[sprseries$Yr == 2023]
   Bratio_min <- min(sprseries$Deplete, na.rm = TRUE)
   Bratio_minyr <- sprseries$Yr[sprseries$Deplete %in% Bratio_min]
   Bratio_max_recent <- max(sprseries$Deplete[sprseries$Yr %in% 2010:2020])
   Bratio_maxyr_recent <- sprseries$Yr[sprseries$Deplete %in% Bratio_max_recent]
   
   # now round values (didn't do above to have matches to find years)
   smry_equ <- format(round(smry_equ), big.mark=",")
   smry_2023 <- format(round(smry_2023), big.mark=",")
   smry_min <- format(round(smry_min), big.mark=",")
   smry_max_recent <- format(round(smry_max_recent), big.mark=",")

   Bratio_2023 <- round(Bratio_2023, 3)
   Bratio_min <- round(Bratio_min, 3)
   Bratio_max_recent <- round(Bratio_max_recent, 3)

```

The base model estimate of biomass time series is similar to previous assessments with estimated biomass of ages 3 and older estimated around `r smry_equ` mt in the unfished equilibrium, declining to a minimum of `r smry_min` mt in `r smry_minyr`, rebuilding quickly to a recent peak of `r smry_max_recent` mt in `r smry_maxyr_recent` due to 3 years of very high recruitments from 2006 to 2008, and then declining to `r smry_2023` mt in 2023 due to low or average recruitment in the years since.

In terms of fraction of unfished spawning output, the minimum was `r Bratio_min` in `r Bratio_minyr`, the recent peak was `r Bratio_max_recent` in `r Bratio_maxyr_recent` and the 2023 estimate is `r Bratio_2023` (Figure \ref{fig:es-depl}, Table \ref{tab:ssbES}).

\input{tex_tables/b_SSB_ES.tex}

```{r, results = 'asis'}
if (!is.na(mod_base$log_det_hessian)) {
  # with Hessian
  add_figure(
  filein = file.path(mod_loc_relative, "plots",
  "ts7_Spawning_output_with_95_asymptotic_intervals_intervals.png"), 
  caption = "Estimated time series of spawning output (circles and line: median; light broken lines: 95 percent intervals) for the base model",
  label = 'es-ssb')
} else {
  # no uncertainty (model not converged or run with -nohess)
  add_figure(
  filein = file.path(mod_loc_relative, "plots",
  "ts7_Spawning_output.png"), 
  caption = "Estimated time series of spawning output for the base model",
  label = 'es-ssb')
}
```

```{r, results = 'asis'}
if (!is.na(mod_base$log_det_hessian)) {
  # with Hessian
  add_figure(
  filein = file.path(mod_loc_relative, "plots", "ts9_Relative_spawning_output_intervals.png"), 
  caption = "Estimated time series of fraction of unfished spawning output (circles and line: median; light broken lines: 95 percent intervals) for the base model",
  label = 'es-depl')
} else {
  # no uncertainty (model not converged or run with -nohess)
  add_figure(
  filein = file.path(mod_loc_relative, "plots", "ts9_Relative_spawning_output.png"), 
  caption = "Estimated time series of fraction of unfished spawning output for the base model",
  label = 'es-depl')
}
```

\clearpage

## Recruitment{-}

Recruitment dynamics (Figure \ref{fig:es-recruits}, Table \ref{tab:recrES}) are assumed to follow Beverton-Holt stock-recruit function and the steepness parameter was fixed at the value of 0.8, which is the mean of steepness prior probability distribution, derived from meta-analysis of flatfish stocks. The level of virgin recruitment (R0) is estimated to inform the magnitude of the initial stock size. Annual recruitment is treated as stochastic. ‘Main’ recruitment deviations were estimated for modeled years that had information about recruitment, between 1959 and 2020. We additionally estimated ‘early’ deviations between 1845 and 1958. 

The recruitment time series is punctuated by four large recruitment events in 1965, 2006, 2007, and 2008. The 1965 recruitment was estimated at about 40 million age-0 recruits while the latter three were in the 25-30 million range. The rest of the time series is close to 10 million recruits per year. The recruitment in 2012 was almost 20 million (a positive deviation from the stock-recruit curve of 0.56) but the years since are estimated to have had below-average or close-to-average recruitment, with a low point in 2017 of about 8 million recruits (deviation = -0.44).

\input{tex_tables/c_Recr_ES.tex}

```{r, results = 'asis'}
if (!is.na(mod_base$log_det_hessian)) {
  # with Hessian
  add_figure(
  filein = file.path(mod_loc_relative, "plots", "ts11_Age-0_recruits_(1000s)_with_95_asymptotic_intervals.png"), 
  caption = "Estimated time series of age-0 recruits (1000s) for the base model with 95 percent intervals",
  label = 'es-recruits')
} else {
  # no uncertainty (model not converged or run with -nohess)
  add_figure(
  filein = file.path(mod_loc_relative, "plots", "ts11_Age-0_recruits_(1000s).png"), 
  caption = "Estimated time series of age-0 recruits (1000s) for the base model",
  label = 'es-recruits')
}
```

```{r, results = 'asis'}
if (!is.na(mod_base$log_det_hessian)) {
  # with Hessian
  add_figure(
  filein = file.path(mod_loc_relative, "plots", "recdevs2_withbars.png"), 
  caption = "Estimated time series of recruitment deviations",
  label = 'es-rec-devs')
} else {
  # no uncertainty (model not converged or run with -nohess)
  add_figure(
  filein = file.path(mod_loc_relative, "plots", "recdevs1_points.png"), 
  caption = "Estimated time series of recruitment deviations",
  label = 'es-rec-devs')
}
```

\clearpage
## Exploitation status{-}

Two measures of exploitation are fishing intensity and exploitation rate. Fishing intensity is defined here as 1 - SPR, where SPR is the equilibrium spawning output at a given combination of F and selectivity relative to spawning output at unfished equilibrium. Using the units of 1-SPR means that more intense fishing is associated with a higher value. The value of 1-SPR in the absence of fishing is 0 and the maximum is 1.0 if all spawning fish are being killed before spawning. The \gls{pfmc} has chosen an SPR target of 0.3 for `r spp` so harvest which leads to SPR below 0.3, or fishing intensity (1-SPR) greater than 0.7 would be overfishing. Exploitation rate is defined as the catch relative to age 3+ biomass. This metric is included because interpretation is simple, but it is not used as a basis for management. 

The estimated time series of exploitation (Figures \ref{fig:es-1-spr} and \ref{fig:es-phase}, Table \ref{tab:exploitES}) shows an accelerating increase in fishing intensity and exploitation rate with a peak around 1990 when the 1-SPR increased to 0.91 and the exploitation rate was close to 0.4. These rates are estimated to have declined slowly up to the point where the overfishing declaration led to dramatic decrease in catch in 2010, when 1-SPR fell below the reference point to about 0.64. The fishing intensity has increased since that time due to the rebuilding of the stock, but is estimated to have remained below the reference point in the years since.

\input{tex_tables/d_SPR_ES.tex}

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc_relative, "plots", "SPR3_ratiointerval.png"), 
caption = "Estimated time series of the fishing intensity (1 - SPR), where SPR is the spawning potential ratio, with approximate 95% asymptotic intervals. The horizontal line at 0.7 corresponds to SPR = 0.3, the management reference point for petrale sole. The horizontal line at 1.0 corresponds to SPR = 0 (all spawning fish removed from the population).",
label = 'es-1-spr')
```

\clearpage

## Ecosystem considerations{-}

Haltuch et al. [-@haltuch_oceanographic_2020] examined the relationship between `r spp` recruitment and oceanographic drivers based on model output from a Regional Ocean Modeling System (ROMS) model for the California Current Ecosystem (Neveu et al. 2016). The results suggested that ROMS output might be useful as the basis for an environmental index of recruitment for `r spp`, to allow for better model precision and near-term forecasting. However, the ROMS model used by Haltuch et al. [-@haltuch_oceanographic_2020] was consistent in structure and inputs for 1980-2010. In 2011, the model was updated, and from 2011 forward, ROMS outputs exhibit distinct discontinuities with 1980-2010 period, showing changes in scale and trend across the 2010/2011 boundary for multiple drivers used to inform the index by Haltuch et al. [-@haltuch_oceanographic_2020]. Appendix A of this report illustrates the discontinuities in ROMS products over the 2010/2011 boundary and discusses potential reasons for why the ROMS time series changed after the model was updated in 2011.

Due to discontinuity in ROMS models between periods before and after 2011, the index developed by Haltuch et al. [-@haltuch_oceanographic_2020] could not be used for this stock assessment. However, an alternative oceanographic model products by Copernicus Marine Environment Monitoring Service (CMEMS) (https://marine.copernicus.eu/) and Mercator Ocean International (MOI) (https://www.mercator-ocean.eu/) were investigated to test if this modelling framework could be used to produce an environmental index of `r spp` recruitment. This new effort is ongoing, and has not yet been published or reviewed by the \gls{ssc} of the \gls{pfmc}. Appendix A of this report describes the most recent, preliminary efforts in developing a new environmental index of `r spp` recruitment based on CMEMS products. In this assessment, we explore the impact of the current CMEMS-based environmental index of `r spp` recruitment via sensitivity analysis.

 

## Reference points{-}

Unfished spawning stock output for `r spp` is estimated to be 22.9 trillion eggs (95% confidence interval: 18-28 trillion eggs). The management biomass target for `r spp` is defined as 25% of the unfished spawning output ($B_{25\%}$), which is estimated by the model to be 5.7 trillion eggs (95% confidence interval: 4.5-6.9 trillion eggs), which corresponds, in a theoretical equilibrium state, to an exploitation rate (catch / age 3+ biomass) of 0.18 (Figure \ref{fig:es-yield}, Table \ref{tab:referenceES}). This harvest rate provides an equilibrium yield of 2481 mt at $B_{25\%}$ (95% confidence interval: 2120-2841 mt). Catch limits are determined by an SPR = 30% reference point which is associated with equilibrium exploitation rate of 0.17. The model estimate of maximum sustainable yield (MSY) is 2482 mt (95% confidence interval: 2121-2842 mt). The estimated spawning stock output at MSY is 5.5 trillion eggs (95% confidence interval: 4.3-6.7 trillion eggs). The exploitation rate corresponding to the estimated  $F_{MSY}$ proxy of SPR = 29% is 0.18. 

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc_relative, "plots", "SPR4_phase.png"), 
caption = "Phase plot of estimated 1-SPR versus fraction unfished for the base model.",
label = 'es-phase')
```

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc_relative, "plots", "yield2_yield_curve_with_refpoints.png"), 
caption = "Equilibrium yield curve for the base case model. Values are based on the most recent
fishery selectivities and retention curves and with steepness fixed at 0.80.",
label = 'es-yield')
```

\clearpage

\input{tex_tables/e_ReferencePoints_ES_FIXED.tex}

\clearpage

## Management performance{-}

In the last ten years total dead catches (including estimated dead discards) 
have been below the annual catch limit, although the attainment has been greater than
90\% in many of those years.

```{r, tables-vi, results = 'asis'}
table_vi <- read.csv("../tables/table_vi.csv")
sa4ss::table_format(x = table_vi,
                 caption = "Recent trend in the overfishing limits (OFLs), the acceptable biological catches (ABCs), the annual catch limits (ACLs), the total landings, and total mortality (mt). The 'Total dead' values may differ from Groundfish Expanded Mortality Multi-Year (GEMM) estimates because the discard mortality is computed within the model based on estimated retention curves.",
                 label = 'table-vi'
                 )
```

## Unresolved problems and major uncertainties{-}

Uncertainty in this assessment model is explicitly captured in the asymptotic confidence intervals reported throughout this assessment for key parameters and management quantities. These intervals reflect the uncertainty in the model fit to the data sources included in the assessment, but do not include uncertainty associated with alternative model configurations or fixed parameters. To explore uncertainty associated with alternative model configurations and evaluate the responsiveness of model outputs to changes in key model assumptions, a variety of sensitivity runs were performed, including runs with different assumptions model structure and treatment of data, life-history parameters, stock-recruitment parameters, and many others. The uncertainty in natural mortality, stock-recruit steepness and the unfished recruitment level was also explored through likelihood profile analysis. Additionally, a retrospective analysis was conducted where the model was run after successively removing data from recent years, one year at a time.

Main life history parameters, such as natural mortality and stock-recruit curve steepness, generally contribute significant uncertainty to stock assessments. These quantities are essential for understanding the dynamics of the stock and determining projected rebuilding. Steepness in this assessment was fixed at 0.8, which is meta-analytical steepness prior for Pleuronectidae. When estimated, steepness was approaching the upper parameter bound of 1, which was considered unrealistic as it was associated with less plausible estimates natural mortality around 0.10 (compared to base model estimates of 0.14 for females). Steepness likelihood profile illustrated that the starting and ending biomass and associated fraction unfished show almost no change across a wide range of steepness values. Natural mortality was estimated for both sexes using meta-analytical prior, but likelihood profile showed that the starting and ending biomass as well as associated fraction unfished are more sensitive to the changes in natural mortality than in steepness. In past several assessments, natural mortality was used as major axis of uncertainty. 

## Decision table and projections{-}

Not required in draft assessment undergoing review.

## Scientific uncertainty{-}

The model estimated uncertainty around the `r endyr+1` spawning output for is $\sigma$ = `r round(mod_base$Pstar_sigma, 2)`. The uncertainty around the OFL is $\sigma$ = `r round(mod_base$OFL_sigma, 2)`. These values are lower than for many West Coast groundfish stocks for several reasons: large `r spp` sample sizes of length and age data from fisheries and surveys, high frequency of occurrence in the \gls{s-wcgbt} thanks to petrale primarily residing in trawlable habitat within the scope of the survey, and strong contrast in the data caused by fishing down the stock to a low level followed by rapid rebuilding. Nevertheless, these $\sigma$ values surely underestimate the overall uncertainty as they do not incorporate the model structural uncertainty and do not account for any time-varying dynamics other than recruitment.

The estimated uncertainty values are lower than the Category 1 default $\sigma = 0.5$, so all projections will use the default $\sigma$.

## Research and data needs{-}

Progress on a number of research topics and data issues would substantially improve the ability of this assessment to reliably and precisely model `r spp` population dynamics in the future:

\begin{enumerate}

\item Continue research toward better understanding of how climate forcing impacts density-independent survival during `r spp` early life stages and further development of environmental recruitment index. Such index can provide additional information on recruitment, not captured by other sources, in most recent years when youngest cohorts may be not yet selected by either surveys or fisheries. 

\item The extent of spatial and temporal variability on productivity processes such as growth, recruitment, and maturity is currently unknown and would benefit from further research. It would allow to better understand patterns we see in data and account for potential spatio-temporal variability in life history parameters in the model. 

\item Studies on stock structure and movement of `r spp` indicating transboundary movement of `r spp` between U.S. and Canadian waters. Within the scope of this assessment, we explored including multiple data sources from British Columbia waters to the base model via sensitivity analysis. Further studies of transboundary movement of `r spp` between U.S. and Canadian waters would be beneficial for understanding the extent of `r spp` movements and to help lay the groundwork for future collaborative effort between U.S. and Canada and potential transboundary assessment.

\item The analytical solution for catchability (i.e., observed / predicted biomass) for the \gls{s-wcgbt} is above 1.0 in the base model.  This was also the case in previous assessments of `r spp` and other flatfish assessments.  Further research into flatfish habitat use and behaviors in response to survey gear will enhance the interpretation of catchability values for `r spp` off the West Coast.

\end{enumerate}
