\clearpage
# Figures
```{r, read-csv, results = "asis"}
plot_info = read.csv(file.path(mod_loc_relative, "plots", "plotinfotable_for_doc.csv"))
```

\clearpage
<!-- don't allow section headers to float relative to figure locations -->
<!-- \floatplacement{figure}{H}  -->

## Data 
<!-- ====================================================================== --> 
<!-- *******************    Assessment Map      *************************** --> 
<!-- ====================================================================== --> 


```{r, data-plot, results = "asis"}
sa4ss::add_figure(
filein = file.path(mod_loc_relative, "plots", "data_plot2.png"), 
caption = 
  "Data presence by year for each fleet, where circle area is
relative within a data type. Circles are proportional to
total catch for catches; to precision for indices, discards, and
mean body weight observations; and to total sample size for
compositions.",
label = "data-plot"
)
```

<!-- ====================================================================== -->  
<!-- ****************** Catches Used in the Model ************************* --> 
<!-- ====================================================================== -->  

## Catch and estimated discards

```{r, catch-figures-r4ss, results = "asis"}
sa4ss::add_figure(
filein = file.path(mod_loc_relative, "plots", "catch2 landings stacked.png"), 
caption = "Landings (mt) by fleet used in the base model.",
label = "catch-figures-r4ss"
)

sa4ss::add_figure(
filein = file.path(mod_loc_relative, "plots", "catch16 landings + dead discards.png"), 
caption = "Landings plus dead discards (mt) by fleet as estimated in the base model.",
label = "catchdiscard-figures-r4ss"
)

```

<!-- ====================================================================== --> 
<!-- ******************* Data Used in the Model *************************** --> 
<!-- ====================================================================== --> 


## Biology


## Model Results

### Bridging


### Model Structure

\clearpage

### Estimated Biology
```{r, results = 'asis'}
sa4ss::add_figure(
filein = file.path(mod_loc_relative, "plots", "bio1_sizeatage.png"),
caption = "Model estimated length-at-age in the beginning of the year. Shaded area indicates 95 percent distribution of length-at-age around the estimated growth curve",
label = 'mod-est-len-age')

sa4ss::add_figure(
filein = file.path(mod_loc_relative, "plots", "bio3_sizeatage_plus_WT_and_MAT.png"),
caption = "Relationship between growth, maturity, and weight. Length at age is in the top-left panel with weight (thick lines) and maturity (thin lines) shown in top-right and lower-left panels",
label = 'mod-est-len-age2')

sa4ss::add_figure(
filein = file.path("../figures/Petrale_fxn_ogive.png"),
caption = "Maturity ogive showing fit to observed fraction mature within each length bin.",
label = 'maturity')
```


### Selectivity
```{r, results = 'asis'}
sa4ss::add_figure(
filein = file.path(mod_loc_relative, "plots",  "sel01_multiple_fleets_length1.png"), 
caption = paste("Ending-year selectivity at length for multiple fleets.",
  "Solid lines are female selectivity, dashed are male."),
label = 'sel_length')

sa4ss::add_figure(
filein = file.path(mod_loc_relative, "plots",  "sel02_multiple_fleets_age2.png"), 
caption = paste("Ending-year selectivity at age derived from selectivity at length",
  "(solid female, dashed male)."),
label = 'sel-age')

sa4ss::add_figure(
filein = file.path(mod_loc_relative, "custom_plots", "selectivity_time-varying_fem.png"), 
caption = paste("Time-varying female selectivity (top) and retention (bottom)",
  "for the fishing fleets. Retention is the same for females and males.",
  "Note: the legend shows the "),
label = 'sel-tv-fem')
sa4ss::add_figure(
filein = file.path(mod_loc_relative, "custom_plots",  "selectivity_time-varying_mal.png"), 
caption = paste("Time-varying male selectivity (top) and retention (bottom)",
  "for the fishing fleets. Retention is the same for females and males."),
label = 'sel-tv-mal')

```


### Recruitment

```{r, results = 'asis'}
sa4ss::add_figure(
filein = file.path(mod_loc_relative, "plots",  "recdevs2_withbars.png"), 
caption = paste("Recruitment deviations with 95% intervals.",
  "Black points and lines represent the 'main' period while the blue",
  "points and lines are the early and late periods."),
label = 'recdevs')

sa4ss::add_figure(
filein = file.path(mod_loc_relative, "custom_plots",  "recruit_fit_bias_adjust.png"), 
caption = paste0("Bias adjustment applied to the recruitment deviations (red line). ",
  "Points are transformed variances relative to $\\sigma_R = ", mod_base$sigma_R_in, "$."),
label = 'biasadj')
```
\clearpage


### Fits to Data

\clearpage

#### Fits to indices

```{r, results = 'asis'}
sa4ss::add_figure(
filein = file.path(mod_loc_relative, "plots",  "index2_cpuefit_WCGBTS.png"), 
caption = paste("Fit to index data for WCGBTS. Lines indicate 95% uncertainty",
  "interval around index values based on the model assumption of lognormal error."),
label = 'index-wcgbts-fit')

sa4ss::add_figure(
filein = file.path(mod_loc_relative, "plots",  "index2_cpuefit_Triennial.png"), 
caption = paste("Fit to index data for the Triennial survey. Lines indicate 95% uncertainty",
  "interval around index values based on the model assumption of lognormal error.",
  "Thicker lines indicate input uncertainty before addition of",
  "estimated additional uncertainty parameter."),
label = 'index-tri-fit')
```

\clearpage

#### Fits to Length and Age Compositions
<!-- TODO: add Francis plots -->
```{r, results = 'asis'}
sa4ss::add_figure(
filein = file.path(mod_loc_relative, "plots",  "comp_lenfit__aggregated_across_time.png"), 
caption = paste("Length composition aggregated across years by fleet with the",
  "model estimated fit to the data by sex (red female and blue male)."),
label = 'len-agg-fit')

sa4ss::add_figure(
filein = file.path(mod_loc_relative, "plots",  "comp_lendat__multi-fleet_comparison.png"), 
caption = paste("Length composition data for all fleets",
  "(red female, blue male, grey unsexed)."),
label = 'len-all-bubbles')

sa4ss::add_figure(
filein = file.path(mod_loc_relative, "plots",  "comp_lenfit__multi-fleet_comparison.png"), 
caption = paste("Pearson residuals for fit to length composition data for all fleets",
  "(red female, blue male, grey unsexed).",
  "Closed bubbles are positive residuals (observed > expected) and open bubbles",
  "are negative residuals (observed < expected)."),
label = 'len-all-pearson')

sa4ss::add_figure(
filein = file.path(mod_loc_relative, "plots",  "comp_agedat__multi-fleet_comparison.png"), 
caption = paste("Age composition data for all fleets",
  "(red female, blue male)."),
label = 'age-all-bubbles')

sa4ss::add_figure(
filein = file.path(mod_loc_relative, "plots",  "comp_agefit__multi-fleet_comparison.png"), 
caption = paste("Pearson residuals for fit to age composition data for all fleets",
  "(red female, blue male).",
  "Closed bubbles are positive residuals (observed > expected) and open bubbles",
  "are negative residuals (observed < expected)."),
label = 'age-all-pearson')

sa4ss::add_figure(
filein = file.path(mod_loc_relative, "plots", "comp_agefit__aggregated_across_time.png"), 
caption = "Age composition data aggregated across time by fleet and sex (red female and blue male).",
label = 'agg-marg-age-fit')

sa4ss::add_figure(
filein = file.path(mod_loc_relative, "plots", "comp_condAALdat_bubflt4mkt0_page1.png"), 
caption = paste("Conditional age-at-length data from WCGBTS",
  "(data plot 1 of 2, interleaved with the residual plots to facilitate flipping back and forth)."),
label = 'caal-dat1')

sa4ss::add_figure(
filein = file.path(mod_loc_relative, "plots", "comp_condAALfit_residsflt4mkt0_page1.png"), 
caption = paste("Pearson residuals for conditional age-at-length data from WCGBTS",
  "(residual plot 1 of 2, interleaved with the data plots to facilitate flipping back and forth)."),
label = 'caal-fit1')

sa4ss::add_figure(
filein = file.path(mod_loc_relative, "plots", "comp_condAALdat_bubflt4mkt0_page2.png"), 
caption = paste("Conditional age-at-length data from WCGBTS",
  "(data plot 2 of 2, interleaved with the residual plots to facilitate flipping back and forth)."),
label = 'caal-dat2')

sa4ss::add_figure(
filein = file.path(mod_loc_relative, "plots", "comp_condAALfit_residsflt4mkt0_page2.png"), 
caption = paste("Pearson residuals for conditional age-at-length data from WCGBTS",
  "(residual plot 2 of 2, interleaved with the data plots to facilitate flipping back and forth)."),
label = 'caal-fit2')

```

\clearpage

#### Fits to discard fractions and mean weight observations

```{r, results = 'asis'}
sa4ss::add_figure(
filein = file.path(mod_loc_relative, "plots",  "discard_fitNorth.png"), 
caption = paste("Discard fraction for the North fishery with 95% intervals (black) with fit of expected value (blue)."),
label = 'discard_fitNorth')

sa4ss::add_figure(
filein = file.path(mod_loc_relative, "plots",  "discard_fitSouth.png"), 
caption = paste("Discard fraction for the South fishery with 95% intervals (black) with fit of expected value (blue)."),
label = 'discard_fitSouth')

sa4ss::add_figure(
filein = file.path(mod_loc_relative, "plots",  "bodywt_fit_fltNorth.png"), 
caption = paste("Mean individual body weight of the discards for the North fishery with 95% intervals (black) with fit of expected value (blue)."),
label = 'mnwt_fitNorth')

sa4ss::add_figure(
filein = file.path(mod_loc_relative, "plots",  "bodywt_fit_fltSouth.png"), 
caption = paste("Mean individual body weight of the discards for the South fishery with 95% intervals (black) with fit of expected value (blue)."),
label = 'mnwt_fitSouth')

```

\clearpage

### Time-series

<!-- ======================================================= -->  
<!-- ****************** Time Series ************************ --> 
<!-- ======================================================= -->
```{r, results = 'asis'}
sa4ss::add_figure(
filein = file.path("../figures/historical_assessment_timeseries.png"), 
caption = paste("Comparison of biomass time series (colored lines) across recent assessments.",
  "Units are estimated biomass of females and males ages 3 and older.",
  "Total mortality from the base model (black bars) is included as well.",
  "Spawning biomass is not comparable across these assessments because the inclusion",
  "of a fecundity relationship starting in 2023 which results in spawning output in units of eggs."),
label = 'smry-bio-historical')


```{r, results = 'asis'}
sa4ss::add_figure(
filein = file.path(mod_loc_relative, "plots", "ts4_Summary_biomass_(mt).png"), 
caption = "Estimated biomass of females and males ages 3 and older.",
label = 'smry-bio')

sa4ss::add_figure(
filein = file.path(mod_loc_relative, "plots", "ts7_Spawning_output_with_95_asymptotic_intervals_intervals.png"), 
caption = "Estimated time series of female spawning output with approximate 95% asymptotic intervals.",
label = 'ssb')

sa4ss::add_figure(
filein = file.path(mod_loc_relative, "plots", "ts9_Relative_spawning_output_intervals.png"), 
caption = "Estimated time series of relative spawning output with approximate 95% asymptotic intervals.",
label = 'depl')


sa4ss::add_figure(
filein = file.path(mod_loc_relative, "plots", "SPR3_ratiointerval.png"), 
caption = "Estimated time series of the relative fishing intensity with approximate 95% asymptotic intervals.",
label = 'spr')

info <- plot_info %>% dplyr::filter(grepl("SPR4_phase", label))
sa4ss::add_figure(
filein = file.path(mod_loc_relative, "plots", "SPR4_phase.png"), 
caption = info$caption,
label = 'phase')

```



\clearpage

### Sensitivity Analyses and Retrospectives
```{r, results = 'asis'}
sa4ss::add_figure(
filein = file.path(mod_loc_relative, "custom_plots", "sens_timeseries_index1.png"), 
label = 'sens-index1',
caption = paste("Time series of spawning biomass (top) and fraction of unfished",
            "(bottom) for the sensitivity analyses related to",
            "index data."))
```

```{r, results = 'asis'}
sa4ss::add_figure(
filein = file.path(mod_loc_relative, "custom_plots", "sens_timeseries_comp1.png"), 
label = 'sens-comp1',
caption = paste("Time series of spawning biomass (top) and fraction of unfished",
            "(bottom) for the sensitivity analyses related to",
            "composition data."))

sa4ss::add_figure(
filein = file.path(mod_loc_relative, "custom_plots", "sens_timeseries_comp2.png"), 
label = 'sens-comp2',
caption = paste("Time series of recruitment (top) and recruitment deviations",
            "(bottom) for the sensitivity analyses related to",
            "composition data."))

sa4ss::add_figure(
filein = file.path(mod_loc_relative, "custom_plots", "sens_timeseries_sel1.png"), 
label = 'sens-sel1',
caption = paste("Time series of spawning biomass (top) and fraction of unfished",
            "(bottom) for the sensitivity analyses related to",
            "selectivity assumptions."))

sa4ss::add_figure(
filein = file.path(mod_loc_relative, "custom_plots", "sens_timeseries_bio1.png"), 
label = 'sens-bio1',
caption = paste("Time series of spawning biomass (top) and fraction of unfished",
            "(bottom) for the sensitivity analyses related to",
            "biology."))

# recruitment sensitivities
sa4ss::add_figure(
filein = file.path(mod_loc_relative, "custom_plots", "sens_timeseries_recruit1.png"), 
label = 'sens-recruit1',
caption = paste("Time series of spawning biomass (top) and fraction of unfished",
            "(bottom) for the sensitivity analyses related to",
            "recruitment."))
sa4ss::add_figure(
filein = file.path(mod_loc_relative, "custom_plots", "sens_timeseries_recruit2.png"), 
label = 'sens-recruit2',
caption = paste("Time series of recruitment (top) and recruitment deviations",
            "(bottom) for the sensitivity analyses related to",
            "recruitment."))

```

```{r, results = 'asis'}
# define path to retro and profile figs
figures_diags <- "../figures/diags_model34/"
```

```{r, results = 'asis'}
add_figure(
filein = file.path(figures_diags, "retro_compare2_spawnbio_uncertainty.png"), 
caption = "Change in the estimate of spawning output when the most recent 5 years of data area removed sequentially",
label = 'retro-ssb')
```

\pagebreak

```{r, results = 'asis'}
add_figure(
filein = file.path(figures_diags, "retro_compare13_indices_flt4.png"), 
caption = "Change in the fit to the WCGBTS index when the most recent 5 years of data area removed sequentially",
label = 'retro-index')
```

\pagebreak

```{r, results = 'asis'}
add_figure(
filein = file.path(figures_diags, "retro_compare4_Bratio_uncertainty.png"),
caption = "Change in the estimate of fraction unfished when the most recent 5 years of data area removed sequentially",
label = 'retro-depl')
```

### Likelihood Profiles

```{r, results = 'asis'}
add_figure(
filein = file.path(figures_diags, "piner_panel_SR_LN(R0).png"), 
caption = "Change in the negative log-likelihood across a range of log(R~0~) values",
label = 'r0-profile')
```

```{r, results = 'asis'}
add_figure(
filein = file.path(figures_diags, "parameter_panel_SR_LN(R0).png"), 
caption = paste("Change in quantities of interest related to spawning output",
  "across a range of log(R~0~) values:",
  "Fraction of unfished spawning output in 2023 (top-right),",
  "Spawning output in 2023 (bottom-right), and ",
  "Unfished equilibrium spawning output (bottom-left).",
  "These are shown along with the change in total negative log-likelihood",
  "(top-left, matches previous figure)."),
label = 'r0-profile-pars')
```


\pagebreak

```{r, results = 'asis'}
add_figure(
filein = file.path(figures_diags, "piner_panel_SR_BH_steep.png"), 
caption = "Change in the negative log-likelihood across a range of steepness (h) values",
label = 'h-profile')
```
```{r, results = 'asis'}
add_figure(
filein = file.path(figures_diags, "parameter_panel_SR_BH_steep.png"), 
caption = paste("Change in quantities of interest related to spawning output",
  "across a range of steepness (h) values:",
  "Fraction of unfished spawning output in 2023 (top-right),",
  "Spawning output in 2023 (bottom-right), and ",
  "Unfished equilibrium spawning output (bottom-left).",
  "These are shown along with the change in total negative log-likelihood",
  "(top-left, matches previous figure)."),
label = 'h-profile-pars')
```

```{r, results = 'asis'}
add_figure(
filein = file.path(figures_diags, "piner_panel_NatM_uniform_Fem_GP_1.png"), 
caption = "Change in the negative log-likelihood across a range of female natural mortality (M) values",
label = 'm-profile')
```

```{r, results = 'asis'}
add_figure(
filein = file.path(figures_diags, "parameter_panel_NatM_uniform_Fem_GP_1.png"), 
caption = paste("Change in quantities of interest related to spawning output",
  "across a range of female natural mortality (M) values:",
  "Fraction of unfished spawning output in 2023 (top-right),",
  "Spawning output in 2023 (bottom-right), and ",
  "Unfished equilibrium spawning output (bottom-left).",
  "These are shown along with the change in total negative log-likelihood",
  "(top-left, matches previous figure)."),
label = 'm-profile-pars')
```


### Reference Points and Forecasts