---
geometry: margin=1in
month: "`r format(Sys.Date(), '%B')`"
year: "`r format(Sys.Date(), '%Y')`"
preamble: |
output:
  sa4ss::techreport_pdf:
    default
  bookdown::pdf_document2:
    keep_tex: true
lang: en
papersize: letter
csl: canadian-journal-of-fisheries-and-aquatic-sciences.csl
---

```{r opts, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
knitr::knit_hooks$set(plot = function(x,options) {
      base = knitr::opts_knit$get('base.url')
      if (is.null(base)) base = ''
      alt = ifelse (is.null(options$alt),"",options$alt)
      cap = ifelse (is.null(options$caption),"",options$caption)
      if (alt != ""){
        sprintf('![%s](%s%s "%s")', cap, base, x, alt)
      } else {
        sprintf('![%s](%s%s)', cap, base, x)
        }
  })

load("00opts.Rdata")

if(file.exists("00mod.Rdata")){
  load("00mod.Rdata")
} else {
  print("Model output not being read. Please run the read_model() function.")
}

```

<!--chapter:end:00a.Rmd-->

---
author:
  - name: Ian G. Taylor
    code: 1
    first: I
    middle: G
    family: Taylor
  - name: Vladlena Gertseva
    code: 1
    first: V
    middle: ''
    family: Gertseva
author_list: Taylor, I.G., V. Gertseva
affiliation:
  - code: 1
    address: Northwest Fisheries Science Center, U.S. Department of Commerce, National
      Oceanic and Atmospheric Administration, National Marine Fisheries Service, 2725
      Montlake Boulevard East, Seattle, Washington 98112
address: ^1^Northwest Fisheries Science Center, U.S. Department of Commerce, National
  Oceanic and Atmospheric Administration, National Marine Fisheries Service, 2725
  Montlake Boulevard East, Seattle, Washington 98112
---

<!--chapter:end:00authors.Rmd-->

---
bibliography:
  - sa4ss.bib
---

<!--chapter:end:00bibliography.Rmd-->

---
title: |
  Status of petrale sole (_Eopsetta jordani_) \
  along the US West coast in 2023
---

<!--chapter:end:00title.Rmd-->

\pagebreak
\pagenumbering{roman}
\setcounter{page}{1}

\renewcommand{\thetable}{\roman{table}}
\renewcommand{\thefigure}{\roman{figure}}


\setlength\parskip{0.5em plus 0.1em minus 0.2em}

<!--chapter:end:01a.Rmd-->

```{r executive, echo = FALSE}
executive <- list()
executive[["stock"]] <- paste0("This assessment reports the status of ",
  spp, " (_", spp.sci, "_) off the ", coast, 
  " coast using data through 2022.")
```

# Executive summary{-}

## Stock{-}

`r executive[["stock"]]`

## Catches{-}

Adding text here for example.
trends and current levels.
Include Table for last 10 years.
Include Figure with long-term estimates.

\clearpage

\input{tex_tables/a_Catches_ES.tex}

```{r, results = 'asis'}
add_figure(
filein = file.path("..", mod_loc, "plots", "catch2 landings stacked.png"), 
caption = "Landings by fleet used in the base model where catches in metric tons by fleet are stacked",
label = 'es-catch')
```

\clearpage

## Data and assessment{-}

This assessment uses the stock assessment framework
Stock Synthesis 
```{r es-modelversion, echo=FALSE} 
if (exists("model")) {
  strsplit(model$SS_version,";")[[1]][1]
}
```
(SS3).

Replace text with
date of last assessment,
type of assessment model,
data available,
new information, and
information lacking.

## Stock biomass and dynamics{-}

Replace text with
trends and current levels relative to virgin or historic levels and
description of uncertainty.
Include Table for last 10 years.
Include Figure with long-term estimates.

\input{tex_tables/b_SSB_ES.tex}

```{r, results = 'asis'}
if (!is.na(mod_base$log_det_hessian)) {
  # with Hessian
  add_figure(
  filein = file.path("..", mod_loc, "plots",
  "ts7_Spawning_output_with_95_asymptotic_intervals_intervals.png"), 
  caption = "Estimated time series of spawning output (circles and line: median; light broken lines: 95 percent intervals) for the base model",
  label = 'es-ssb')
} else {
  # no uncertainty (model not converged or run with -nohess)
  add_figure(
  filein = file.path("..", mod_loc, "plots",
  "ts7_Spawning_output.png"), 
  caption = "Estimated time series of spawning output for the base model",
  label = 'es-ssb')
}
```

```{r, results = 'asis'}
if (!is.na(mod_base$log_det_hessian)) {
  # with Hessian
  add_figure(
  filein = file.path("..", mod_loc, "plots", "ts9_Relative_spawning_output_intervals.png"), 
  caption = "Estimated time series of fraction of unfished spawning output (circles and line: median; light broken lines: 95 percent intervals) for the base model",
  label = 'es-depl')
} else {
  # no uncertainty (model not converged or run with -nohess)
  add_figure(
  filein = file.path("..", mod_loc, "plots", "ts9_Relative_spawning_output.png"), 
  caption = "Estimated time series of fraction of unfished spawning output for the base model",
  label = 'es-depl')
}
```

\clearpage

## Recruitment{-}

Replace text with
trends and current levels relative to virgin or historic levels and
description of uncertainty.
Include Table for last 10 years.
Include Figure with long-term estimates.


\input{tex_tables/c_Recr_ES.tex}

```{r, results = 'asis'}
if (!is.na(mod_base$log_det_hessian)) {
  # with Hessian
  add_figure(
  filein = file.path("..", mod_loc, "plots", "ts11_Age-0_recruits_(1000s)_with_95_asymptotic_intervals.png"), 
  caption = "Estimated time series of age-0 recruits (1000s) for the base model with 95 percent intervals",
  label = 'es-recruits')
} else {
  # no uncertainty (model not converged or run with -nohess)
  add_figure(
  filein = file.path("..", mod_loc, "plots", "ts11_Age-0_recruits_(1000s).png"), 
  caption = "Estimated time series of age-0 recruits (1000s) for the base model",
  label = 'es-recruits')
}
```

```{r, results = 'asis'}
if (!is.na(mod_base$log_det_hessian)) {
  # with Hessian
  add_figure(
  filein = file.path("..", mod_loc, "plots", "recdevs2_withbars.png"), 
  caption = "Estimated time series of recruitment deviations",
  label = 'es-rec-devs')
} else {
  # no uncertainty (model not converged or run with -nohess)
  add_figure(
  filein = file.path("..", mod_loc, "plots", "recdevs1_points.png"), 
  caption = "Estimated time series of recruitment deviations",
  label = 'es-rec-devs')
}
```

\clearpage
## Exploitation status{-}

Replace text with
total catch divided by exploitable biomass or SPR harvest rate.
Include Table for last 10 years.
Include Figure with trend in f relative to target vs. trend in biomass relative to the target.

\input{tex_tables/d_SPR_ES.tex}

```{r, results = 'asis'}
add_figure(
filein = file.path("..", mod_loc, "plots", "SPR2_minusSPRseries.png"), 
caption = "Estimated 1 - relative spawning ratio (SPR) by year for the base model. The management target is plotted as a red horizontal line and values above this reflect harvest in excess of the proxy harvest rate",
label = 'es-1-spr')
```

\clearpage

## Ecosystem considerations{-}

Replace text with
a summary of reviewed environmental and ecosystem factors that appear to be correlated with stock dynamics.
These may include variability in they physical environment, habitat, competitors, prey, or predators that directly or indirectly affects the stock's status, vital rates (growth, survival, productivity/recruitment) or range and distribution.
Note which, if any, ecosystem factors are used in the assessment and how (e.g., as background information, in data preparations, as data inputs, in decisions about model structure).

## Reference points{-}

Replace text with
management targets and definition of overfishing, including the harvest rate that brings the stock to equilibrium at $B_{40\%}$, i.e., the $B_{MSY}$ proxy and the equilibrium stock size that results from fishing at the default harvest rate, i.e., the $F_{MSY}$ proxy.
Include Table of estimated reference points for ssb, SPR, exploitation rate, and yield based on SSB proxy for MSY, SPR proxy for MSY, and estimated MSY values.

```{r, results = 'asis'}
add_figure(
filein = file.path("..", mod_loc, "plots", "SPR4_phase.png"), 
caption = "Phase plot of estimated 1-SPR versus fraction unfished for the base model",
label = 'es-phase')
```

```{r, results = 'asis'}
add_figure(
filein = file.path("..", mod_loc, "plots", "yield2_yield_curve_with_refpoints.png"), 
caption = "Equilibrium yield curve for the base case model. Values are based on the 2020
fishery selectivities and with steepness fixed at 0.80",
label = 'es-yield')
```

\input{tex_tables/e_ReferencePoints_ES.tex}

\clearpage

## Management performance{-}

Include Table of most recent 10 years of
catches in comparison with OFL, ABC, HG, and OY/ACL values,
overfishing levels,
actual catch and discard.
Include OFL (encountered), OFL (retained), and OFL (dead) if different due to discard and discard mortality.

## Unresolved problems and major uncertainties{-}

Replace text with
any special issues that complicate scientific assessment, questions about the best model scenario, etc.

## Decision table and projections{-}

Replace text with
projected yields (OFL, ABC, and ACL), spawning biomass, and stock depletion levels for each year.
OFL calculations should be based on the assumption that future catches equal ABCs and not OFLs.

## Scientific uncertainty{-}

Replace text with
the sigma value and the basis for its calculation.

## Research and data needs{-}

Replace text with
information gaps that seriously impede the stock assessment.

<!--chapter:end:01executive.Rmd-->

\pagebreak
\setlength{\parskip}{5mm plus1mm minus1mm}
\pagenumbering{arabic}
\setcounter{page}{1}
\renewcommand{\thefigure}{\arabic{figure}}
\renewcommand{\thetable}{\arabic{table}}
\setcounter{table}{0}
\setcounter{figure}{0}

<!--chapter:end:10a.Rmd-->

# Introduction

This updated assessment does not attempt to reiterate all background information for `r spp` presented in the 2013 assessment document. Instead, only a few key assumptions are restated, along with a detailed description of changes made during the course of the update. Those interested in a more complete description of `r spp` life-history and the details of previous assessments should refer to the 2013 assessment [@haltuch_status_2013].
  
## Basic Information
`r Spp` (\emph{`r spp.sci`}) is a right-eyed flounder in the family Pleuronectidae ranging from the western Gulf of Alaska to the Coronado Islands, northern Baja California [@kramer_guide_1995; @love_milton_resource_2005] with a preference for soft substrates at depths ranging from 0-550 m [@love_milton_resource_2005]. Common names include brill, California sole, Jordan's flounder, cape sole, round nose sole, English sole, soglia, petorau, nameta, and tsubame garei [@smith_report_1937; @gates_designated_1974; @love_milton_probably_1996; @eschmeyer_field_1983]. In northern and central California `r spp` are dominant on the middle and outer continental shelf. PacFIN fishery logbook data show that adults are caught in depths from 18 to 1,280 m off the U.S. West Coast with a majority of the catches of `r spp` being taken between 70-220 m during March through October, and between 290-440 m during November through February.

There is little information regarding the stock structure of `r spp` off the U.S. Pacific coast. No genetic research has been undertaken for `r spp` and there is no other published research indicating separate stocks of `r spp` within U.S. waters. Tagging studies show adult `r spp` can move up to 350 - 390 miles, having the ability to be highly migratory with the possibility for homing ability [@alverson_results_1957]. Juveniles show little coast-wide or bathymetric movement while studies suggest that adults generally move inshore and northward onto the continental shelf during the spring and summer to feeding grounds and offshore and southward during the fall and winter to deep water spawning grounds [@horton_species_1989; @love_milton_probably_1996]. Adult `r spp` can tolerate a wide range of bottom temperatures [@perry_environmental_1994].

Tagging studies indicate some mixing of adults between different spawning groups. DiDonato and Pasquale [-@didonato_migration_1970] reported that five fish tagged on the Willapa Deep grounds during the spawning season were recaptured during subsequent spawning seasons at other deepwater spawning grounds, as far south as Eureka (northern California) and the Umpqua River (southern Oregon). However, Pedersen [-@pedersen_movements_1975] reported that most of the fish (97%) recaptured from spawning grounds in winter were originally caught and tagged on those same grounds.

Mixing of fish from multiple deep water spawning grounds likely occurs during the spring and summer when `r spp` are feeding on the continental shelf. Fish that were captured, tagged, and released off the northwest Coast of Washington during May and September were subsequently recaptured during winter from spawning grounds off Vancouver Island (British Columbia, 1 fish), Heceta Bank (central Oregon, 2 fish), Eureka (northern California, 2 fish), and Halfmoon Bay (central California, 2 fish) (Pederson, 1975). Fish tagged south of Fort Bragg (central California) during July 1964 were later recaptured off Oregon (11 fish), Washington (6 fish), and Swiftsure Bank (southwestern tip of Vancouver Island, 1 fish) (D. Thomas, California Department of Fish and Game, Menlo Park, CA, cited by Sampson and Lee [-@sampson_assessment_1999]).

The highest densities of spawning adults off of British Columbia, as well as of eggs, larvae and juveniles, are found in the waters around Vancouver Island. Adults may utilize nearshore areas as summer feeding grounds and non-migrating adults may stay there during winter [@starr_petrale_2004].

Past assessments completed by Demory [-@demory_progress_1984,], Turnock et al. [-@turnock_status_1993], and Sampson and Lee [-@sampson_assessment_1999] considered `r spp` in the Columbia and U.S.-Vancouver INPFC areas a single stock. Sampson and Lee (1999) assumed that `r spp` in the Eureka and Monterey INPFC areas represented two additional distinct socks. The 2005 `r spp` assessment assumed two stocks, northern (U.S.-Vancouver and Columbia INPFC areas) and southern (Eureka, Monterey and Conception INPFC areas), to maintain continuity with previous assessments. Three stocks (West Coast Vancouver Island, Queen Charlotte Sound, and Heceta Strait) are considered for `r spp` in the waters off British Columbia, Canada [@starr_petrale_2004]. The 2009, 2011, 2013, and 2015 assessments integrate the previously separate north-south assessments to provide a coast-wide status evaluation. The decision to conduct a single-area assessment is based on strong evidence of a mixed stock from tagging studies, a lack of genetic studies on stock structure, and a lack of evidence for differences in growth between the 2005 northern and southern assessment areas and from examination of the fishery size-at-age data, as well as confounding differences in data collection between Washington, Oregon, and California. This 2019 update assessment provides a coast-wide status evaluation for `r spp` using data through 2018.

Fishing fleets are separated both geographically and seasonally to account for spatial and seasonal patterns in catch given the coast-wide assessment area. The `r spp` fisheries possess a distinct seasonality, with catches peaking during the winter months, so the fisheries are divided into winter (November-February) and summer (March-October) fisheries. Note that the "fishing year" for this assessment (November 1 to October 31) differs from the standard calendar year. The U.S.-Canadian border is the northern boundary for the assessed stock, although the basis for this choice is due to political and current management needs rather than the population dynamics. Given the lack of clear information regarding the status of distinct biological populations, this assessment treats the U.S. `r Spp` resource from the Mexican border to the Canadian border as a single coast-wide stock.


## Life History
`r Spp` spawn during the winter at several discrete deepwater sites (270-460 m) off the U.S. West Coast, from November to April, with peak spawning taking place from December to February [@harry_analysis_1956; @best_petrale_1960; @gregory_validity_1976; @castillo_g.c._environmental_1993; @reilly_recreational_1994; @castillo_latitudinal_1995; @love_milton_probably_1996]. Females spawn once each year and fecundity varies with fish size, with one large female laying as many as 1.5 million eggs [@porter_notes_1964]. `r Spp` eggs are planktonic, ranging in size from 1.2 to 1.3 mm, and are found in deep water habitats at water temperatures of 4-10 degrees C and salinities of 25-30 ppt [@best_petrale_1960; @ketchen_population_1966; @alderdice_effects_1971; @gregory_validity_1976]. The duration of the egg stage can range from approximately 6 to 14 days [@alderdice_effects_1971; @love_milton_probably_1996]. The most favorable conditions for egg incubation and larval growth are 6-7 degrees C and 27.5-29.5 ppt [@ketchen_population_1966; @alderdice_effects_1971]. 

`r Spp` spawn during the winter at several discrete deepwater sites (270-460 m) off the U.S. West Coast, from November to April, with peak spawning taking place from December to February [@harry_time_1959; @best_petrale_1960; @gregory_validity_1976; @castillo_g.c._environmental_1993; @reilly_recreational_1994; @love_milton_probably_1996]. Females spawn once each year and fecundity varies with fish size, with one large female laying as many as 1.5 million eggs [@porter_notes_1964]. `r Spp` eggs are planktonic, ranging in size from 1.2 to 1.3 mm, and are found in deep water habitats at water temperatures of 4-10 degrees C and salinities of 25-30 ppt [@best_petrale_1960; @ketchen_population_1966;  @alderdice_effects_1971; @gregory_validity_1976]. The duration of the egg stage can range from approximately 6 to 14 days [@alderdice_effects_1971; @love_milton_probably_1996]. The most favorable conditions for egg incubation and larval growth are 6-7 degrees C and 27.5-29.5 ppt [@ketchen_population_1966; @alderdice_effects_1971; @castillo_latitudinal_1995].

Adult `r spp` achieve a maximum size of around 50 cm and 63 cm for males and females, respectively [@best_e.a._movements_1963; @pedersen_movements_1975]. The maximum length reported for `r spp` is 70 cm [@eschmeyer_field_1983; @love_milton_resource_2005] while the maximum observed break-and-burn age is 31 years [@haltuch_status_2013].

## Ecosystem Considerations
`r Spp` juveniles are carnivorous, foraging on annelid worms, clams, brittle star, mysids, sculpin, amphipods, and other juvenile flatfish [@casilla_essential_1998; @pearsall_diet_2007]. Predators on juvenile `r spp` include adult `r spp` as well as other larger fish [@casilla_essential_1998] while adults are preyed upon by marine mammals, sharks, and larger fishes [@trumble_abundance_1995; @love_milton_probably_1996; @casilla_essential_1998].

One of the ambushing flatfishes, adult `r spp` have diverse diets that become more piscivorous at larger sizes [@allen_ecology_2006]. Adult `r spp` are found on sandy and sand-mud bottoms [@eschmeyer_field_1983] foraging for a variety of invertebrates including, crab, octopi, squid, euphausiids, and shrimp, as well as anchovies. hake, herring, sand lance, and other smaller rockfish and flatfish [@kravitz_food_1977; @birtwell_fish_1984; @reilly_recreational_1994; @love_milton_probably_1996; @pearsall_diet_2007]. In Canadian waters evidence suggests that `r spp` tend to prefer herring [@pearsall_diet_2007]. On the continental shelf `r spp` generally co-occur with English sole, rex sole, Pacific sanddab, and rock sole [@kravitz_food_1977].

Ecosystem factors have not been explicitly modeled in this assessment, but there are several aspects of the California current ecosystem that may impact `r spp` population dynamics and warrant further research. Castillo [-@castillo_g.c._fluctuations_1992] and Castillo et al. [-@castillo_latitudinal_1995] suggest that density-independent survival of early life stages is low and show that offshore Ekman transportation of eggs and larvae may be an important source of variation in year-class strength in the Columbia INPFC area. The effects of the Pacific Decadal Oscillation (PDO) on California current temperature and productivity [@mantua_pacific_1997] may also contribute to non-stationary recruitment dynamics for `r spp`. The prevalence of a strong late 1990s year-class for many West Coast groundfish species suggests that environmentally driven recruitment variation may be correlated among species with relatively diverse life history strategies. Although current research efforts along these lines are limited, a more explicit exploration of ecosystem processes may be possible in future `r spp` stock assessments.

## Historical and Current Fishery Information
`r Spp` have been caught in the flatfish fishery off the U.S. Pacific coast since the late 19th century. The fishery first developed off of California where, prior to 1876, fishing in San Francisco Bay was by hand or set lines and beach seining [@scofield_trawling_1948]. By 1880 two San Francisco based trawler companies were running a total of six boats, extending the fishing grounds beyond the Golden Gate Bridge northward to Point Reyes [@scofield_trawling_1948]. Steam trawlers entered the fishery during 1888 and 1889, and four steam tugs based out of San Francisco were sufficient to flood market with flatfish [@scofield_trawling_1948]. By 1915 San Francisco and Santa Cruz trawlers were operating at depths of about 45-100 m with catches averaging 10,000 lbs per tow or 3,000 lbs per hour [@scofield_trawling_1948]. Flatfish comprised approximately 90% of the catch with 20-25% being discarded as unmarketable [@scofield_trawling_1948]. During 1915 laws were enacted that prohibited dragging in California waters and making it illegal to possess a trawl net from Santa Barbara County southward [@scofield_trawling_1948]. By 1934 twenty 56-72 foot diesel engine trawlers operated out of San Francisco fishing between about 55 and 185 m [@scofield_trawling_1948]. From 1944-1947 the number of California trawlers fluctuated between 16 and 46 boats [@scofield_trawling_1948]. Although the flatfish fishery in California was well developed by the 1950s and 1960s, catch statistics were not reported until 1970 [@heimann_pacific_1970]. In this early California report `r spp` landings during 1916 to 1930 were not separated from the total flatfish landings. 

The earliest trawl fishing off Oregon began during 1884-1885, and the fishery was solidly established by 1937, with the fishery increasing rapidly during WWII (Harry and Morgan, 1961). Initially trawlers stayed close to the fishing grounds adjacent to Newport and Astoria, operating at about 35-90 m between Stonewall Bank and Depoe Bay. Fishing operations gradually extended into deep water. For example, Newport-based trawlers were commonly fishing at about 185 m in 1949, at about 185-365 m by 1952, and at about 550 m by 1953.

Alverson and Chatwin [-@alverson_results_1957] describe the history of the `r spp` fishery off of Washington and British Columbia with fishing grounds ranging from Cape Flattery to Destruction Island. `r Spp` catches off of Washington were small until the late 1930s with the fishery extending to about 365 m following the development of deepwater rockfish fisheries during the 1950s.

By the 1950s the `r spp` fishery was showing signs of depletion with reports suggesting that `r spp` abundance had declined by at least 50% from 1942 to 1947 [@harry_analysis_1956]. Sampson and Lee [-@sampson_assessment_1999] reported that three fishery regulations were implemented during 1957-67: 1) a winter closure off Oregon, Washington and British Columbia, 2) a 3,000 lb per trip limit, and 3) no more than two trips per month during 1957. With the 1977 enactment of the Magnuson Fishery Conservation and Management Act (MFCMA) the large foreign-dominated fishery that had developed since the late 1960s was replaced by the domestic fishery that continues today. `r Spp` are harvested almost exclusively by bottom trawls in the U.S. West Coast groundfish fishery. Recent `r spp` catches exhibit marked seasonal variation, with substantial portions of the annual harvest taken from the spawning grounds during December and January. Evidence suggests that the winter fishery on the deepwater spawning grounds developed sporadically during the 1950s and 1960s as fishers discovered new locations (e.g., @alverson_results_1957; @ketchen_population_1966). Both historical and current `r spp` fisheries have primarily relied upon trawl fleets. Fishery removals were divided among 4 fleets: 1) winter North trawl, 2) summer North trawl, 3) winter South trawl, and 4) summer South trawl. Landings for the North fleet are defined as fish landed in Washington and Oregon ports. Landings for the South fleet are defined as fish landed in California ports.

<!-- commenting out section that references tables and values not yet available -->
<!-- Historical landings reconstructions show peak catches from the summer fishery occurred during the 1940s and 1950s and subsequently declined, during which time the fleet moved to fishing in deeper waters during the winter. After the period of peak landings during the 1940s and 1950s, total landings were somewhat stable until about the late 1970s, and then generally declined until the mid-2000s. -->

## Summary of Management History and Performance
Beginning in 1983 the Pacific Fishery Management Council (PFMC) established coast-wide annual catch limits (ACLs) for the annual harvests of `r spp` in the waters off the U.S. West Coast. The first assessment of West Coast `r spp` occurred in 1984 [@demory_progress_1984]. Based on the 1999 assessment a coast-wide ACL of 2,762 mt was specified and remained unchanged between 2001 and 2006.

The 2005 assessment of `r spp` stock assessment split the stock into two areas, the northern area that included U.S.-Vancouver and Columbia INPFC areas and the southern area that included the Eureka, Monterey and Conception INPFC areas [@lai_stock_2005]. While `r spp` stock structure is not well understood, CPUE and geographical differences between states were used to support the use of two separate assessment areas. In 2005 `r spp` were estimated to be at 34 and 29% of unfished spawning stock biomass in the northern and southern areas, respectively. In spite of different models and data, the biomass trends were qualitatively similar in both areas, providing support for a coast-wide stock. This assessment estimated that `r spp` had historically been below the Pacific Council's minimum stock size threshold of 25% of unfished biomass from the mid-1970s until just prior to the completion of the assessment, with estimated harvest rates in excess of the target fishing mortality rate implemented for `r spp` at that time (F40%). However, the 2005 stock assessment determined that the stock was in the precautionary zone and was not overfished (i.e., the spawning stock biomass was not below 25% of the unfished spawning stock biomass). Based on the 2005 stock assessment results, ACLs were set at 3,025 mt and 2,919 mt for 2007 and 2008, respectively, with an ACT of 2,499 mt for both years. 

In comparison to the 1999 assessment of `r spp`, the 2005 assessment represented a significant change in the perception of `r spp` stock status. The stock assessment conducted in 1999 (Washington-Oregon only) estimated the spawning stock biomass in 1998 at 39% of unfished stock biomass. Although the estimates of 1998 spawning-stock biomass were little changed between the 1999 and 2005 (Northern area) assessments, the estimated depletion in the 2005 assessment was much lower. The change in status between the 1999 and 2005 analyses was due to the introduction of a reconstructed catch history in 2005, which spanned the entire period of removals. The 1999 stock assessment used a catch history that started in 1977, after the bulk of the removals from the fishery had already taken place. Thus the 1999 stock assessment produced a more optimistic view of the petrale stock's level of depletion. The stock's estimated decline in status between the 2005 and 2009 assessments was driven primarily by a significant decline in the trawl-survey index over that period. The 2011 assessment concluded that the stock status continued to be below the target of 25% of unfished biomass.

The 2009 coast-wide stock assessment estimated that the `r spp` stock had declined from its 2005 high to 11.6% of the unfished spawning stock biomass [@haltuch_status_2009]. The `r spp` was declared overfished based on newly adopted management targets (e.g., target spawning biomass for flatfish stocks defined as 25% and overfished threshold of 12.5% of unfished spawning stock biomass) resulting in a rebuilding plan and catch restrictions for `r spp`. The stock was declared rebuilt based on the results of the 2015 update stock assessment which estimated the coastwide biomass at 30.7% of unfished spawning stock output with ACLs of 3,136 and 3,013 in 2017 and 2018 respectively [@stawitz_stock_2015]. 

For additional information on changes in the `r spp` fishery please see the 2013 stock assessment [@haltuch_status_2013].

## Fisheries off Canada and Alaska
The Canadian fishery developed rapidly during the late 1940s to mid-1950s following the discovery of `r spp` spawning aggregations off the West Coast of Vancouver Island [@anon_fish_2001]. Annual landings of `r spp` in British Columbia peaked at 4,800 mt in 1948 but declined significantly after the mid-1960s [@anon_fish_2001]. By the 1970s, analysis conducted by Pederson [-@pedersen_movements_1975] suggested that `r spp` abundance was low and abundance remained low into the 1990s. In the early 1990s vessel trip quotas were established to try to halt the decline in `r spp` abundance [@anon_fish_2001]. Winter quarter landings of `r spp` were limited to 44,000 lb per trip during 1985-91; to 10,000 lb per trip during 1991-95; and to 2,000 lb per trip in 1996. Biological data collected during 1980-1996 showed a prolonged decline in the proportion of young fish entering the population [@anon_fish_2001]. Therefore, no directed fishing for `r spp` has been permitted in Canada since 1996 due to a continuing decline in long term abundance [@fargo_j.j._flatfish_1997; @anon_fish_2001]. As of 2005 `r spp` off of British Columbia were treated as three "stocks" and were still considered to be at low levels. The recent assessments for the Canadian stocks have been based on catch histories and limited biological data. 

In Alaska `r spp` are not targeted in the Bering Sea/Aleutian Island fisheries and are managed as a minor species in the "Other Flatfish" stock complex.

<!--chapter:end:11introduction.Rmd-->

# Data

Data comprise the foundational components of stock assessment models.
The decision to include or exclude particular data sources in an assessment model depends on many factors.
These factors often include, but are not limited to,
the way in which data were collected (e.g., measurement method and consistency);
the spatial and temporal coverage of the data;
the quantity of data available per desired sampling unit;
the representativeness of the data to inform the modeled processes of importance;
timing of when the data were provided;
limitations imposed by the Terms of Reference; and
the presence of an avenue for the inclusion of the data in the assessment model.
Attributes associated with a data source can change through time,
as can the applicability of the data source when different modeling approaches are explored (e.g., stock structure or time-varying processes).
Therefore, the specific data sources included or excluded from this assessment should not necessarily constrain the selection of data sources applicable to future stock assessments for `r spp`.
Even if a data source is not directly used in the stock assessment they can provide valuable insights into biology, fishery behavior, or localized dynamics.

Data from a wide range of programs were available for possible inclusion in the current assessment model.
Descriptions of each data source included in the model (Figure \@ref(fig:data-plot)) and sources that were explored but not included in the base model are provided below.
Data that were excluded from the base model were explicitly explored during the development of this stock assessment or have not changed since their past exploration in a previous `r spp` stock assessment.
In some cases, the inclusion of excluded data sources were explored through sensitivity analyses (see Section \@ref(assessment-model)).

<!--chapter:end:20data.Rmd-->

## Fishery-Dependent Data
Fishery removals were divided among two fleets: 1) North and 2) South. Landings for the North fleet are defined as fish landed in Washington and Oregon ports. Landings for the South feet are defined as fish landed in California ports. The landings of Petrale sole are made primarily by groundfish bottom trawl gear; landings by gear types other than bottom trawl have been inconsequential, averaging less than 2.5% of the coast-wide landings. The non-trawl landings (that consist of only a small fraction of the total landings) are included along with the trawl landings in a single gears fleets.

##Commercial Fishery Landings
All landings for this assessment were summarized by state (based on port of landing). Landings for Washington and Oregon are summed into a single North fleet (consistent with previous assessment) due to the fact that vessels commonly fish and land in each other's waters and ports.

###Recent landings
Recent commercial landings of Petrale sole (1981-2022 for California and Washington; 1987-2022 for Oregon) were obtained from the Pacific Fisheries Information Network (PacFIN), a regional fisheries database that manages fishery-dependent information in cooperation with west coast state agencies and NOAA Fisheries (www.pacfin.com). Catch data were extracted by state and then combined into the fishing fleets used in the assessment. Time series of recent (PacFIN era) landings by state are shown in *Figure XX. 

###Historical landings
Historical landings of Petrale sole were reconstructed by state, by year. The Washington historical landings (1938 -1980) were obtained from Washington Department of Fish and Wildlife, who recently conducted historical catch reconstruction for Petrale sole (pers. comm. T. Tsou and G. Lippert, WDFW). The historical Oregon landings (1932-1986) were obtained from Oregon historical catch reconstruction, conducted by Oregon Department of Fish and Wildlife in collaboration with NWFSC (Karnowski et al. 2014). The California historical landings were informed by several sources. The most recent “historical” period (between 1969 and 1980) were available from the California Cooperative Groundfish Survey (CalCOM) database. Earlier landing records (between 1931 and 1968) were reconstructed by the Southwest Fisheries Science Center (SWFSC; Ralston et al. 2010). Ralston et al. (2010) included only California landings made in California ports, therefore Ralston et al. (2010) estimates were supplemented by catches landed in California ports, but made in waters of Oregon and Washington, provided by the SWFSC (pers. comm. E.J. Dick, SWFSC). The earliest California landings were obtained from California Department of Fish and Game (CDFG) Fish Bulletins for 1916-1930 landings (Heimann and Carlisle 1970) as reconstructed by Lai et al. (2005) and used in previous assessments (REF HERE). The California fishery began in 1876, but no landings data are available from 1876-1915. Therefore, consistent with previous Petrale sole assessments, a linear interpolation between landings of 1 ton in 1876 and the landings recorded for 1916 are used to filling this period.

##Discard
Data on discards of Petrale sole are available from two different data sources. In 2002, the West Coast Groundfish Observer Program (WCGOP) was implemented on the West Coast of the United States, which began with gathering bycatch and discard information for the limited entry trawl and fixed gear fleets. Observer coverage has expanded to include the California halibut trawl, the nearshore fixed gear and pink shrimp trawl fisheries. Since 2011, trawl fisheries have been managed with catch shares under a system of annual individual fishing quotas (IFQs) for the shoreside sector (i.e., vessels delivering to shoreside processors) and harvest cooperatives for the at-sea hake sectors (catcher-processors who catch and process hake at sea; and Motherships, factory processors that take delivery of hake from catcher vessels at sea). A 100 percent at-sea monitoring of catch using observers and electronic monitoring (EM) is required to participate in the trawl catch share fishery.
The historical source (often referred to as the Pikitch data) comes from a study organized by Ellen Pikitch that collected trawl discards from 1985-1987 (Pikitch et al. 1988). The northern and southern boundaries of the study were 48∘42′ N latitude and 42∘60′ N latitude respectively, which is primarily within the Columbia INPFC area (Pikitch et al. 1988, Rogers and Pikitch 1992). Therefore, this source of discard data is only relevant to North fleet. Participation in the study was voluntary and included vessels using bottom, midwater, and shrimp trawl gears. Observers of normal fishing operations on commercial vessels collected the data, estimated the total weight of the catch by tow, and recorded the weight of species retained and discarded in the sample. Results of the Pikitch data were obtained from NWFSC (pers. comm. John Wallace NWFSC). 

##Fishery Length and Age Data
###Fishery landings compositions

The fishery length and age data for North fleet were obtained from PacFIN BDS database and contained data from Oregon (1966-present) and Washington (1955-present). For South fleet length and age data were obtained from PacFIN BDS database (1986-present) and from the California Cooperative Groundfish Survey (CalCOM, 1948-1992), the latter were provided by CDFG (pers.comm. Brenda Erwin, CDFG).

The historical Oregon data for Petrale sole has change substantially since 2015. The state identified that samples collected prior to 1987 were not collected according to the state's standardized sampling protocol and were mistakenly included in PacFIN as random samples
(pers. comm. Ali Whitman, ODFW). These samples likely represent samples that were collected for special projects. Oregon had removed some of these samples for Petrale sole from PacFIN, but not all. To remove the remaining non-standard samples, Oregon PacFIN data were filtered to remove samples prior to 1987. Although these samples were removed from the length and age composition data for this update assessment, future assessments should evaluate these data to determine if they can be included within the assessment through alternative means (e.g., external estimates of length-at-age or as conditional-age-at-length data).

Commercial length-frequency distributions based on the fishing year were developed for each state for which observations were available. For each fleet, the raw observations (compiled from the PacFIN and CalCOM databases) were expanded to the trip level, to account for the relative size of the landing from which the sample was obtained. The expanded length observations were then expanded by the landings in each state for the combined Washington and Oregon fleet. Age frequencies were computed in the same manner, except that age observations for Washington and Oregon were not combined due to ageing error considerations. The length and age composition expansion was conducted using PacFIN.Utilities function publicly available via GitHUB.

Length and age data collected from commercial landings for each fleet are summarized by the number of trips and fish sampled by year (Tables 11 and 12). Figures 33, 34, and 35 show plots of the commercial length and age composition data across time for each fishery fleet.

The initial input sample sizes (Ninput) for length and age frequency distributions by year were calculated as a function of the number of trips and number of fish via the Stewart Method
(Stewart, INPFC, pers.comm.): 

N_input=N_trips+0.138N_fish		when	N_fish/N_trips <44

N_input=7.06N_trips				when	N_fish/N_trips ≥44

The method is based on analysis of the input and model derived effective sample sizes from west coast groundfish stock assessments. A piece-wise linear regression was used to estimate the increase in effective sample size per sample based on fish-per-sample and the maximum effective sample size for large numbers of individual fish.

###Discard length compositions and mean body weight
Discard length composition data were available from both WCGOP and Pikitch study. WCGOP length composition data were not sex specific; Pikitch data (although available by sex) was very limited (32 fish sampled for length), and was included in the assessment for females and males combined. No age data were available for discarded fish.
WCGOP also provided estimates of discard mean body weight, and these data were used in this assessment for each fishing fleets. The mean body weight of discarded fish are shown in Figures 29 - 32. 



<!--chapter:end:21f-.Rmd-->

## Fishery-Independent Data

<!--chapter:end:21s-.Rmd-->

### \acrlong{s-aslope}

The \gls{s-aslope} operated during the months of October to November aboard the R/V _Miller Freeman_.
Partial survey coverage of the US west coast occurred during the years 1988-1996 and complete coverage (north of 34\textdegree 30\textquotesingle S) during the years 1997 and 1999-2001.
Typically, only these four years that are seen as complete surveys are included in assessments.

<!--chapter:end:21s-aslope.Rmd-->

### \acrlong{s-ccfrp}

Since 2007, the \gls{s-ccfrp} has monitored several areas in California to evaluate the performance of \glspl{mpa}
and understand nearshore fish populations
[@Wendt2009; @Starr2015].
In 2017, the survey expanded beyond the four \Gls{mpa}s in central California
(A&ntilde;o Nuevo, Point Lobos, Point Buchon, and Piedras Blancas)
to include the entire California coast.
Fish are collected by volunteer anglers aboard \glspl{cpfv} guided by one of the following academic institutions based on proximity to fishing location:
Humboldt State University;
Bodega Marine Laboratories;
Moss Landing Marine Laboratories;
Cal Poly San Luis Obispo;
University of California, Santa Barbara; and
Scripps Institution of Oceanography.

Surveys consist of fishing with hook-and-line gear for 30-45 minutes within randomly chosen 500 by 500 m grid cells within and outside \glspl{mpa}.
Prior to 2017, all fish were measured for length and release or descended to depth;
since then, some were sampled for otoliths and fin clips.


<!--chapter:end:21s-ccfrp.Rmd-->

### \acrlong{s-tri}

The \gls{s-tri} was first conducted by the \gls{afsc} in 1977, and the survey continued until 2004 [@weinberg_2001_2002].
Its basic design was a series of equally-spaced east-to-west transects across the continential shelf from which searches for tows in a specific depth range were initiated.
The survey design changed slightly over time.
In general, all of the surveys were conducted in the mid summer through early fall.
The 1977 survey was conducted from early July through late September.
The surveys from 1980 through 1989 were conducted from mid-July to late September.
The 1992 survey was conducted from mid July through early October.
The 1995 survey was conducted from early June through late August.
The 1998 survey was conducted from early June through early August.
Finally, the 2001 and 2004 surveys were conducted from May to July.

Haul depths ranged from 91-457 m during the 1977 survey with no hauls shallower than 91 m.
Due to haul performance issues and truncated sampling with respect to depth, the data from 1977 were omitted from this analysis.
The surveys in 1980, 1983, and 1986 covered the US West Coast south to 36.8\textdegree N latitude and a depth range of 55-366 m.
The surveys in 1989 and 1992 covered the same depth range but extended the southern range to 34.5\textdegree N (near Point Conception).
From 1995 through 2004, the surveys covered the depth range 55-500 m and surveyed south to 34.5\textdegree N.
In 2004, the final year of the \gls{s-tri} series, the \gls{nwfsc} \gls{fram} conducted the survey following similar protocols to earlier years.

<!--chapter:end:21s-tri.RMd-->

### \acrlong{s-wcgbt}

The \Gls{s-wcgbt} is based on a random-grid design;
covering the coastal waters from a depth of 55-1,280 m [@bradburn_2003_2011].
This design generally uses four industry-chartered vessels per year assigned to a roughly equal number of randomly selected grid cells and divided into two 'passes' of the coast.
Two vessels fish from north to south during each pass between late May to early October.
This design therefore incorporates both vessel-to-vessel differences in catchability,
as well as variance associated with selecting a relatively small number (approximately 700) of possible cells from a very large set of possible cells spread from the Mexican to the Canadian borders.

<!--chapter:end:21s-wcgbts.Rmd-->

## Biological Data

### Natural Mortality


### Maturation and Fecundity


### Sex Ratio


### Length-Weight Relationship {#sec-biological-data-Length-Weight-relationship}

`r load("../data/weight-length_pars.rda")`
The weight-length relationship for `r spp` was estimated outside of the 
assessment model by fitting biological data to the standard power function, 
$W = aL^b$ using the R function `PacFIN.Utilities::getWLpars()` 
(where $W$ is weight in kilograms and $L$ is fork length in centimeters). 
The function estimates the relationship on a log-log scale and then uses the 
estimated standard deviation of the observed weights around the expected value 
to calculate the median weight at each length from the resulting lognormal 
distribution.

The parameters were estimated using data from the \gls{s-wcgbt}, where 21,704 
fish collected between 2003 and 2022 had both weight and length available 
of which 57.6\% were female, 42.3\% were male, and 0.1\% were unsexed. 
The resulting relationships were
`r sprintf("$W=%1.9f * L^{%1.3f}$", fa, fb)` for females and
`r sprintf("$W=%1.9f * L^{%1.3f}$", ma, mb)` for males. 
These relationships are very similar to those used for the previous assessment 
[@wetzel_status_2019].

### Growth (Length-at-Age)


### Ageing Precision and Bias

<!--chapter:end:22biology.Rmd-->

## Environmental and Ecosystem Data

<!--chapter:end:23enviro.Rmd-->

# Assessment Model


<!--chapter:end:30model.Rmd-->

## Summary of Previous Assessments and Reviews


### History of Modeling Approaches (not required for an update assessment)


### Most Recent STAR Panel and SSC Recommendations (not required for an update assessment)


### Response to Groundfish Subcommittee Requests (not required in draft)

<!--chapter:end:31summary.Rmd-->

## Model Structure and Assumptions


### Model Changes from the Last Assessment (not required for an update assessment)


### Modeling Platform and Structure
General model specifications (e.g., executable version, model structure, definition of fleets and areas)


### Model Parameters
Describe estimated vs. fixed parameters, priors

### Key Assumptions and Structural Choices

<!--chapter:end:32structure.Rmd-->

## Base Model Results


### Parameter Estimates
<!-- Parameters values are shown in Tables \ref{tab:table-pars-base-1}-\ref{tab:table-pars-base-16}. -->

### Fits to the Data

### Population Trajectory

### Reference Points



<!--chapter:end:33results.Rmd-->

## Model Diagnostics
Describe all diagnostics

### Convergence

### Sensitivity Analyses

### Retrospective Analysis

### Likelihood Profiles

### Unresolved Problems and Major Uncertainties

<!--chapter:end:34diagnostics.Rmd-->

# Management 

## Reference Points

## Unresolved Problems and Major Uncertainties

## Harvest Projections and Decision Tables

## Evaluation of Scientific Uncertainty

## Research and Data Needs

<!--chapter:end:40management.Rmd-->

# Acknowledgments
Here are all the mad props!

<!--chapter:end:41acknowledgments.Rmd-->

\clearpage

# References
<!-- If you want to references to appear somewhere before the end, add: -->
<div id="refs"></div>
<!-- where you want it to appear -->
<!-- The following sets the appropriate indentation for the references
  but it cannot be used with bookdown and the make file because it leads
  to a bad pdf.
\noindent
\vspace{-2em}
\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\setlength{\parskip}{8pt}
 -->

<!--chapter:end:49bibliography.Rmd-->

\clearpage
# Tables


<!-- ====================================================================== -->
<!-- model tables -->
<!-- ====================================================================== -->
```{r tables-model, results = "asis"}
cat("\\begin{landscape}")
#table_pars_split(mod_base)
cat("\\end{landscape}")
```

<!--chapter:end:52tables.Rmd-->

\clearpage
# Figures

## Data 
<!-- ====================================================================== --> 
<!-- *******************    Assessment Map      *************************** --> 
<!-- ====================================================================== --> 


```{r, data-plot, results = "asis"}
sa4ss::add_figure(
filein = file.path(mod_loc_relative, "plots", "data_plot2.png"), 
caption = 
  "Data presence by year for each fleet, where circle area is
relative within a data type. Circles are proportional to
total catch for catches; to precision for indices, discards, and
mean body weight observations; and to total sample size for
compositions.",
label = "data-plot"
)
```

<!-- ====================================================================== -->  
<!-- ****************** Catches Used in the Model ************************* --> 
<!-- ====================================================================== -->  

## Catch and estimated discards

```{r, catch-figures-r4ss, results = "asis"}
sa4ss::add_figure(
filein = file.path(mod_loc_relative, "plots", "catch2 landings stacked.png"), 
caption = "Landings (mt) by fleet used in the base model.",
label = "catch-figures-r4ss"
)

sa4ss::add_figure(
filein = file.path(mod_loc_relative, "plots", "catch16 landings + dead discards.png"), 
caption = "Landings plus dead discards (mt) by fleet as estimated in the base model.",
label = "catchdiscard-figures-r4ss"
)

```

<!-- ====================================================================== --> 
<!-- ******************* Data Used in the Model *************************** --> 
<!-- ====================================================================== --> 


## Biology


## Model Results

### Bridging


### Model Structure

\clearpage

### Estimated Biology
```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc_relative, "plots", "bio1_sizeatage.png"),
caption = "Model estimated length-at-age in the beginning of the year. Shaded area indicates 95 percent distribution of length-at-age around the estimated growth curve",
label = 'mod-est-len-age')

add_figure(
filein = file.path(mod_loc_relative, "plots", "bio3_sizeatage_plus_WT_and_MAT.png"),
caption = "Relationship between growth, maturity, and weight. Length at age is in the top-left panel with weight (thick lines) and maturity (thin lines) shown in top-right and lower-left panels",
label = 'mod-est-len-age2')

add_figure(
filein = file.path("../figures/Petrale_fxn_ogive.png"),
caption = "Maturity ogive showing fit to observed fraction mature within each length bin.",
label = 'maturity')
```


### Selectivity
```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc_relative, "plots",  "sel01_multiple_fleets_length1.png"), 
caption = paste("Ending-year selectivity at length for multiple fleets.",
  "Solid lines are female selectivity, dashed are male."),
label = 'sel_length')

add_figure(
filein = file.path(mod_loc_relative, "plots",  "sel02_multiple_fleets_age2.png"), 
caption = paste("Ending-year selectivity at age derived from selectivity at length",
  "(solid female, dashed male)."),
label = 'sel-age')

# add_figure(
# filein = file.path(mod_loc_relative, "custom_plots", "selectivity_time-varying_fem.png"), 
# caption = paste("Time-varying female selectivity (top) and retention (bottom)",
#   "for the fishing fleets. Retention is the same for females and males."),
# label = 'sel-tv-fem')

# add_figure(
# filein = file.path(mod_loc_relative, "custom_plots",  "selectivity_time-varying_mal.png"), 
# caption = paste("Time-varying male selectivity (top) and retention (bottom)",
#   "for the fishing fleets. Retention is the same for females and males."),
# label = 'sel-tv-mal')

```



### Recruitment


### Fits to Data

### Fits to Data
#### Fits to indices
```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc_relative, "plots",  "index2_cpuefit_WCGBTS.png"), 
caption = paste("Fit to index data for WCGBTS. Lines indicate 95% uncertainty",
  "interval around index values based on the model assumption of lognormal error."),
label = 'index-wcgbts-fit')

add_figure(
filein = file.path(mod_loc_relative, "plots",  "index2_cpuefit_Triennial.png"), 
caption = paste("Fit to index data for the Triennial survey. Lines indicate 95% uncertainty",
  "interval around index values based on the model assumption of lognormal error.",
  "Thicker lines indicate input uncertainty before addition of",
  "estimated additional uncertainty parameter."),
label = 'index-tri-fit')
```

\clearpage

#### Fits to Length and Age Compositions
<!-- TODO: add Francis plots -->
```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc_relative, "plots",  "comp_lenfit__aggregated_across_time.png"), 
caption = paste("Length composition aggregated across years by fleet with the",
  "model estimated fit to the data by sex (red female and blue male)."),
label = 'len-agg-fit')

add_figure(
filein = file.path(mod_loc_relative, "plots",  "comp_lendat__multi-fleet_comparison.png"), 
caption = paste("Length composition data for all fleets",
  "(red female, blue male, grey unsexed)."),
label = 'len-all-bubbles')

add_figure(
filein = file.path(mod_loc_relative, "plots",  "comp_lenfit__multi-fleet_comparison.png"), 
caption = paste("Pearson residuals for fit to length composition data for all fleets",
  "(red female, blue male, grey unsexed).",
  "Closed bubbles are positive residuals (observed > expected) and open bubbles",
  "are negative residuals (observed < expected)."),
label = 'len-all-pearson')

add_figure(
filein = file.path(mod_loc_relative, "plots",  "comp_agedat__multi-fleet_comparison.png"), 
caption = paste("Age composition data for all fleets",
  "(red female, blue male)."),
label = 'age-all-bubbles')

add_figure(
filein = file.path(mod_loc_relative, "plots",  "comp_agefit__multi-fleet_comparison.png"), 
caption = paste("Pearson residuals for fit to age composition data for all fleets",
  "(red female, blue male).",
  "Closed bubbles are positive residuals (observed > expected) and open bubbles",
  "are negative residuals (observed < expected)."),
label = 'age-all-pearson')

add_figure(
filein = file.path(mod_loc_relative, "plots", "comp_agefit__aggregated_across_time.png"), 
caption = "Age composition data aggregated across time by fleet and sex (red female and blue male).",
label = 'agg-marg-age-fit')

add_figure(
filein = file.path(mod_loc_relative, "plots", "comp_condAALdat_bubflt4mkt0_page1.png"), 
caption = paste("Conditional age-at-length data from WCGBTS",
  "(data plot 1 of 2, interleaved with the residual plots to facilitate flipping back and forth)."),
label = 'caal-dat1')

add_figure(
filein = file.path(mod_loc_relative, "plots", "comp_condAALfit_residsflt4mkt0_page1.png"), 
caption = paste("Pearson residuals for conditional age-at-length data from WCGBTS",
  "(residual plot 1 of 2, interleaved with the data plots to facilitate flipping back and forth)."),
label = 'caal-fit1')

add_figure(
filein = file.path(mod_loc_relative, "plots", "comp_condAALdat_bubflt4mkt0_page2.png"), 
caption = paste("Conditional age-at-length data from WCGBTS",
  "(data plot 2 of 2, interleaved with the residual plots to facilitate flipping back and forth)."),
label = 'caal-dat2')

add_figure(
filein = file.path(mod_loc_relative, "plots", "comp_condAALfit_residsflt4mkt0_page2.png"), 
caption = paste("Pearson residuals for conditional age-at-length data from WCGBTS",
  "(residual plot 2 of 2, interleaved with the data plots to facilitate flipping back and forth)."),
label = 'caal-fit2')

```

\clearpage

#### Fits to discard fractions and mean weight observations

```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc_relative, "plots",  "discard_fitNorth.png"), 
caption = paste("Discard fraction for the North fishery with 95% intervals (black) with fit of expected value (blue)."),
label = 'discard_fitNorth')

add_figure(
filein = file.path(mod_loc_relative, "plots",  "discard_fitSouth.png"), 
caption = paste("Discard fraction for the South fishery with 95% intervals (black) with fit of expected value (blue)."),
label = 'discard_fitNorth')

add_figure(
filein = file.path(mod_loc_relative, "plots",  "bodywt_fit_fltNorth.png"), 
caption = paste("Mean individual body weight of the discards for the North fishery with 95% intervals (black) with fit of expected value (blue)."),
label = 'mnwt_fitNorth')

add_figure(
filein = file.path(mod_loc_relative, "plots",  "bodywt_fit_fltSouth.png"), 
caption = paste("Mean individual body weight of the discards for the South fishery with 95% intervals (black) with fit of expected value (blue)."),
label = 'mnwt_fitNorth')

```

\clearpage

### Time-series

<!-- ======================================================= -->  
<!-- ****************** Time Series ************************ --> 
<!-- ======================================================= -->


```{r, results = 'asis'}
add_figure(
filein = file.path(mod_loc_relative, "plots", "ts4_Summary_biomass_(mt).png"), 
caption = "Estimated biomass (summary biomass) of males and females ages 3 and older.",
label = 'tot-bio')

add_figure(
filein = file.path(mod_loc_relative, "plots", "ts7_Spawning_output_with_95_asymptotic_intervals_intervals.png"), 
caption = "Estimated time series of female spawning output with approximate 95% asymptotic intervals.",
label = 'ssb')

add_figure(
filein = file.path(mod_loc_relative, "plots", "ts9_Relative_spawning_output_intervals.png"), 
caption = "Estimated time series of relative spawning output with approximate 95% asymptotic intervals.",
label = 'depl')
```

\clearpage

### Sensitivity Analyses and Retrospectives


### Likelihood Profiles


### Reference Points and Forecasts

<!--chapter:end:53figures.Rmd-->

